---
editor_options:
  chunk_output_type: console
  ---

```{r Notes}
# Before loading this dataset, I recoded some of the Q80 responses, the open text where people indicated whether they open carried, concealed carried, etc. That's what the current data set is we're loading

# Shouldn't use weights here unless we are estimate descriptive data within the whole dataset. When we do analyses on just gun-owners, the weighting scheme will be thrown off because we're subsetting the data so much that the weights no longer make sense

# Missing state gun ownership data from DC so models exclude data from DC
```

```{r Load libraries}
# Data import/export
library(haven)
library(MASS)        # Load before tidyverse so dplyr::select takes precedence
library(tidyverse)
library(openxlsx)
library(httr)

# Modeling
library(lme4)
library(lmerTest)
library(emmeans)
library(estimatr)
library(survey)
library(nnet)
library(ordinal)

# Tidying & robust SEs
library(broom)
library(broom.mixed)
library(sandwich)
library(lmtest)
library(clubSandwich)

# Tables & output
library(flextable)
library(officer)
library(modelsummary)
library(glue)

# Misc
library(usmap)
library(audio)

dir.create("output/figures", recursive = TRUE, showWarnings = FALSE)
```

```{r Utility functions}
# Resolve state abbreviations to full names
resolve_state_name <- function(x) {
  full1 <- state.name[match(toupper(x), state.abb)]
  full2 <- ifelse(x %in% state.name, x, NA_character_)
  ifelse(!is.na(full1), full1, full2)
}

# Significance stars for p-values
sig_stars <- function(p) {
  dplyr::case_when(
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    p < .10  ~ "+",
    TRUE     ~ ""
  )
}

# Create matched sample with complete cases for all model variables
make_matched_sample <- function(dv, data, individual_controls, state_sets) {
  all_state_covs <- unique(unlist(state_sets))
  rhs <- paste("cc_no_permit",
               individual_controls,
               paste(all_state_covs, collapse = " + "),
               sep = " + ")

  vars <- all.vars(as.formula(paste(dv, "~", rhs)))
  vars <- unique(c(vars, "state"))

  missing_vars <- setdiff(vars, names(data))
  if (length(missing_vars)) {
    stop("These variables are missing from `data`: ",
         paste(missing_vars, collapse = ", "))
  }

  n0 <- nrow(data)
  out <- data %>% dplyr::filter(dplyr::if_all(dplyr::all_of(vars), ~ !is.na(.)))
  n1 <- nrow(out)

  message("Matched-sample rows: ", n1, " (from ", n0, "). ",
          "Dropped ", n0 - n1, " rows with any missing among: ",
          paste(vars, collapse = ", "))

  out
}

# Fit survey-weighted model and extract ORs with stars
fit_and_extract_svy <- function(formula, data) {
  des <- svydesign(
    ids = ~state,
    weights = ~wgtpi,
    data = data
  )

  m <- svyglm(formula = formula,
              design = des,
              family = quasibinomial())

  out <- tidy(m) %>%
    mutate(
      OR = exp(estimate),
      coefstars = paste0(sprintf("%.2f", OR), sig_stars(p.value))
    ) %>%
    dplyr::select(term, coefstars)

  list(model = m, table = out)
}

# Fit GLM with cluster-robust SEs and extract ORs with stars
fit_and_extract_crse <- function(formula, data) {
  m <- glm(formula, data = data, family = binomial())

  vc <- sandwich::vcovCL(m, cluster = data$state)

  coef_names <- names(coef(m))
  keep <- coef_names != "(Intercept)"
  vc_reduced <- vc[keep, keep, drop = FALSE]

  SE_CR <- sqrt(diag(vc_reduced))

  td <- broom::tidy(m) %>%
    filter(term != "(Intercept)") %>%
    mutate(
      SE_CR = SE_CR,
      OR    = exp(estimate),
      p_CR  = 2 * pnorm(-abs(estimate / SE_CR)),
      coef_stars = paste0(sprintf("%.2f", OR), sig_stars(p_CR))
    ) %>%
    dplyr::select(term, coef_stars)

  list(model = m, table = td)
}

# Fit mixed-effects logistic model (glmer) and extract ORs with stars
fit_and_extract <- function(formula, data) {
  m <- glmer(
    formula = formula,
    data = data,
    family = binomial(),
    control = glmerControl(optimizer = "bobyqa",
                           optCtrl = list(maxfun = 2e5))
  )

  td <- tidy(m, effects = "fixed") %>%
    mutate(
      OR = exp(estimate),
      coefstars = paste0(sprintf("%.2f", OR), sig_stars(p.value))
    ) %>%
    dplyr::select(term, coefstars)

  list(model = m, table = td)
}

# Fit cumulative link mixed model (clmm) and extract coefficients with stars
fit_and_extract_clmm <- function(formula, data) {
  m <- clmm(
    formula = formula,
    data = data,
    link = "logit",
    Hess = TRUE
  )

  td <- tidy(m, effects = "fixed") %>%
    filter(!term %in% "(Intercept)") %>%
    mutate(
      coefstars = paste0(sprintf("%.2f", estimate), sig_stars(p.value))
    ) %>%
    dplyr::select(term, coefstars)

  list(model = m, table = td)
}
```

```{r Load data}

# load data
dat <- read_csv("data/data_weighted_stata.csv")
#dat <- dat %>% select(-IPAddress, -(ResponseId:LocationLongitude))
```

```{r Clean Zip Codes}

str(dat$zip)
# see how many characters we typically have
dat %>%
  mutate(zip_length = nchar(zip)) %>%
  count(zip_length, name = "n_zipcodes") %>%
  arrange(zip_length)
# about 1,035 that aren't 5 digits
# 1,032 are 4 digits

# see any zip codes that aren't 5 numeric characters
dat %>%
  filter(nchar(zip) != 5 | !str_detect(zip, "^\\d{5}$")) %>%
  count(zip, sort = TRUE)
# Did the 4-digit ones just lose leading 0s?

# are others missing leading zeros?
dat %>%
  filter(str_detect(zip, "^0")) %>%
  count()
# no, so the two 4-digit ones are likely just missing the leading zeros; can change manually; the only plausible zip codes they would be are ones with a leading zero, not with a trailing zero

# go ahead and trim any white space
dat <- dat %>%
  mutate(zip = str_trim(zip)) #trimming any white space but not adding or removing any 0s

# Recode. one had an "O" instead of a zero.
dat <- dat %>%
  mutate(zip = case_when(
    zip == "O3839" ~ "03839",
    TRUE ~ zip
  ))

# replace leading zeros that got lost in the conversion
dat <- dat %>%
  mutate(
    zip = str_trim(as.character(zip)),  # make sure it's character and trim whitespace
    zip = if_else(
      str_detect(zip, "^\\d{1,4}$"),    # only pad ZIPs with fewer than 5 digits
      str_pad(zip, width = 5, side = "left", pad = "0"),
      zip                               # leave valid 5-digit and non-numeric ZIPs alone
    )
  )

# see the character numbers now
dat %>%
  mutate(zip_length = nchar(zip)) %>%
  count(zip_length, name = "n_zipcodes") %>%
  arrange(zip_length)
# the single 3-digit zip code got padded into 5 digits too, with 2 leading zeros

```

```{r Load and Clean HUD Crosswalk Key}

# Get data from the HUD USPS ZIP Code Crosswalk Files API

key <- "YOUR_HUD_API_KEY_HERE"  # Get your free API key at https://www.huduser.gov/hudapi/public/register
url <- "https://www.huduser.gov/hudapi/public/usps"

# Note that type is set to 1 which will return values for the ZIP to Tract file and query is set to All which will return Zip Codes in all states
response <- httr::GET(url, query = list(type = 1, query = "All"), add_headers(Authorization = paste("Bearer", key)))

#check for errors
httr::http_error(response)

#access up the output
output <- httr::content(response)

# Extract the results list
results_list <- output$data$results

# Convert list of named lists into a dataframe
hud_key <- map_dfr(results_list, as_tibble)

# see how many characters we typically have
hud_key %>%
  mutate(zip_length = nchar(zip)) %>%
  count(zip_length, name = "n_zipcodes") %>%
  arrange(zip_length)
# they're all 5 numbers

# trim any white space
hud_key <- hud_key %>%
  mutate(zip = str_trim(zip)) #trimming any white space but not adding or removing any 0s

# right now the key has more than one row per zip code because it's mapping to tract, not just city, state, etc.
hud_zipkey <- hud_key %>%
  distinct(zip, state)  # make it one row per ZIP code

```

```{r Crosswalk with HUD Key}

# join the hud data based on the zip code
dat_merged <- dat %>%
  left_join(hud_zipkey, by = "zip")

# Check to see if it worked
dat_merged %>%
  mutate(match_found = !is.na(state)) %>%
  count(match_found)
# only 35 were unmatched

# quick look at the data
dat_merged %>%
  dplyr::select(zip, state) %>%
  filter(!is.na(state)) %>%
  slice_sample(n = 10)

# See which zips didn't match
dat_merged %>%
  filter(is.na(state)) %>%
  count(zip) %>%
  arrange(desc(n)) %>% 
  print(n = Inf)

# do the unmatched zips exist in the key?
setdiff(dat_merged$zip[is.na(dat_merged$state)], hud_zipkey$zip)
# no

# I didn't get zip codes from multiple years right? Only one version?
output$data$year
output$data$quarter

# create a zip flag column for any that didn't match before overriding some manually
dat_merged <- dat_merged %>%
  mutate(zip_flag = if_else(is.na(state), 1L, 0L))

```

```{r Manually replace unmatched zip codes and remove the rest}

# Check the specific unmatched ones
dat_merged %>%
  filter(is.na(state)) %>%
  count(zip) %>%
  arrange(desc(n)) %>% 
  print(n = Inf)

# 21275 = MD - BOTH
# 19255 = PA (IRS?) - BOTH
# 02266 = inactive; (no longer active from USPS nor find my rep, now organized under 02116); maybe MA now
# 08546 = inactive - BOTH
# 18025 = inactive, maybe PA - BOTH
# 21928 = inactive - BOTH
# 27757 = inactive - BOTH
# 28283 = inactive - BOTH
# 28288 = NC - inactive on find my rep..
# 32337 = FL - BOTH
# 44290 = inactive - BOTH
# 46296 = inactive - BOTH
# 55478 = inactive - BOTH
# 58074 = ND - BOTH
# 60209 = inactive - BOTH
# 75962 = TX (Stephen F Austin University) - BOTH
# 78713 = TX - BOTH
# 83100 = inactive - BOTH
# 86714 = inactive - BOTH
# 88901 = inactive - BOTH
# 90189 = CA - BOTH
# 90410 = CA - BOTH
# 94142 = inactive in USPS but CA in find my rep
# 94230 = inactive in USPS but CA in find my rep
# 94268 = inactive in USPS but CA in find my rep
# 94288 = inactive in USPS but CA in find my rep
# 94299 = inactive - BOTH
# 94394 = inactive - BOTH
# 94529 = inactive in USPS but CA in find my rep
# 99950 = inactive in USPS but AK in find my rep
# M4G 2L2 = so that's Ontario

dat_merged <- dat_merged %>%
  mutate(state = case_when(
    zip == "21275" ~ "MD",
    zip == "19255" ~ "PA",
    zip == "75962" ~ "TX",
    zip == "78713" ~ "TX",
    zip == "32337" ~ "FL",
    zip == "58074" ~ "ND",
    zip == "90189" ~ "CA",
    zip == "90410" ~ "CA",
    zip == "94142" ~ "CA",
    zip == "94230" ~ "CA",
    zip == "94268" ~ "CA",
    zip == "94288" ~ "CA",
    zip == "94529" ~ "CA",
    zip == "99950" ~ "AK",
    TRUE ~ state  # keep existing value for all others
  ))

# filter out rows where state is NA
dat_merged <- dat_merged %>% 
  filter(!is.na(state))

# check to make sure no more unmatched rows
dat_merged %>%
  filter(is.na(state)) %>%
  count(zip) %>%
  arrange(desc(n)) %>% 
  print(n = Inf)

# Remove the one person from Puerto Rico cause we don't have state permissiveness data for that
dat_merged <- dat_merged %>% 
  filter(!state == "PR")
```

```{r Add state-level variables}

# crime data
dat_crime <- read_csv("data/state_crime_2024.csv")
colnames(dat_crime)
str(dat_crime)
# Data regarding crime and solve rates come from the FBI’s Uniform Crime Reporting Program. Violent crime includes homicide, aggravated assault, robbery, and rape. Crime rates are calculated by dividing the number of estimated incidents by the total population for a given geography. Solve rates are the number of reported crimes cleared by arrest or by exceptional means divided by the number of reported crimes.
# from https://projects.csgjusticecenter.org/tools-for-states-to-address-crime/50-state-crime-data/

# trump vote share
dat_trump <- read_csv("data/state_vote_share.csv")
colnames(dat_trump)
str(dat_trump)
# calculated from official state-level returns compiled by the Federal Election Commission and publicly reported election dashboards

# black population
dat_black <- read_csv("data/state_black_population_2020.csv")
colnames(dat_black)
str(dat_black)
# from census 2020

# poverty
dat_poverty <- read_csv("data/state_poverty_2024.csv")
colnames(dat_poverty)
str(dat_poverty)
# from census via https://worldpopulationreview.com/state-rankings/poverty-rate-by-state and https://data.census.gov

# gun ownership
dat_gun_ownership <- read_csv("data/state_gun_ownership_2022.csv")
colnames(dat_gun_ownership)
str(dat_gun_ownership)
# from https://www.rand.org/pubs/tools/TL354.html via https://worldpopulationreview.com/state-rankings/gun-ownership-by-state


# start merging


# helper: map full state names -> abbreviations (incl. DC)
states_full <- c(state.name, "District of Columbia")
states_abb  <- c(state.abb,  "DC")

to_abbrev <- function(x) {
  # x can be full names ("Alabama") or already abbreviations ("AL")
  out <- states_abb[match(x, states_full)]
  # if not a full name, maybe it's already an abbreviation
  out2 <- ifelse(x %in% states_abb, x, NA_character_)
  ifelse(!is.na(out), out, out2)
}

# 1) CRIME (violent_rate_2024)
crime_join <- dat_crime %>%
  mutate(state_abbrev = to_abbrev(state)) %>%
  filter(!is.na(state_abbrev)) %>%
  transmute(
    state = state_abbrev,
    violent_rate_2024
  )

# 2) TRUMP (trump_perc -> numeric)
trump_join <- dat_trump %>%
  rename(state_full = STATE) %>%
  mutate(
    state_abbrev = to_abbrev(state_full),
    trump_perc = as.numeric(str_replace(trump_perc, "%", ""))  # e.g., "52.2%" -> 52.2
  ) %>%
  filter(!is.na(state_abbrev)) %>%
  transmute(
    state = state_abbrev,
    trump_perc_2024 = trump_perc
  )

# 3) BLACK POP SHARE (pct_black_2020)
black_join <- dat_black %>%
  mutate(state_abbrev = to_abbrev(state)) %>%
  filter(!is.na(state_abbrev)) %>%
  transmute(
    state = state_abbrev,
    pct_black_2020
  )

# 4) POVERTY (pct_poverty_2024)
poverty_join <- dat_poverty %>%
  mutate(state_abbrev = to_abbrev(state)) %>%
  filter(!is.na(state_abbrev)) %>%
  transmute(
    state = state_abbrev,
    pct_poverty_2024
  )

# 5) GUN OWNERSHIP (pct_households_gun_own_2022)
gunown_join <- dat_gun_ownership %>%
  mutate(state_abbrev = to_abbrev(state)) %>%
  filter(!is.na(state_abbrev)) %>%
  transmute(
    state = state_abbrev,
    pct_households_gun_own_2022
  )

# Merge into dat_merged
dat_merged <- dat_merged %>%
  left_join(crime_join,  by = "state") %>%
  left_join(trump_join,  by = "state") %>%
  left_join(black_join,  by = "state") %>%
  left_join(poverty_join,by = "state") %>%
  left_join(gunown_join, by = "state")

# Sanity checks
# Any states in the survey not found in the context files?
missing_any <- function(context_df) {
  anti_join(dat_merged %>% distinct(state), context_df, by = "state")
}
missing_any(crime_join)    # expect 0 rows
missing_any(trump_join)
missing_any(black_join)
missing_any(poverty_join)
missing_any(gunown_join)

# Look at new columns
dplyr::glimpse(dat_merged %>% 
                 dplyr::select(state, violent_rate_2024, trump_perc_2024, 
                        pct_black_2020, pct_poverty_2024, pct_households_gun_own_2022) %>% 
                 arrange(state))


# z scored versions for modelling
dat_merged <- dat_merged %>%
  mutate(
    state_violent_rate_2024_z = scale(violent_rate_2024)[,1],
    state_trump_perc_2024_z = scale(trump_perc_2024)[,1],
    state_pct_black_2020_z = scale(pct_black_2020)[,1],
    state_pct_poverty_2024_z = scale(pct_poverty_2024)[,1],
    state_gun_own_2022_z = scale(pct_households_gun_own_2022)[,1]
  )
```

```{r Clean and recode other columns}
# race
sum(is.na(dat_merged$Q5))
sum(is.na(dat_merged$Q4))
# no missing race data

# recode race variable, despite Alexandra's prior coding, which currently leaves lots of missing race data
dat_merged <- dat_merged %>%
  mutate(
    # Hispanic overrides race for dummy coding
    hisp = if_else(Q4 == 1, 1, 0),

    # Non-Hispanic White
    white = case_when(
      hisp == 0 & Q5 == 1 ~ 1,
      hisp == 0 & !is.na(Q5) ~ 0,
      TRUE ~ 0  # if Q5 missing or hisp==1, set to 0
    ),

    # Non-Hispanic Black
    black = case_when(
      hisp == 0 & Q5 == 2 ~ 1,
      hisp == 0 & !is.na(Q5) ~ 0,
      TRUE ~ 0
    ),

    # Other race (non-Hispanic)
    other_race = case_when(
      hisp == 0 & Q5 %in% c(3,4,5,6) ~ 1,
      hisp == 0 & !is.na(Q5) ~ 0,
      TRUE ~ 0
    )
  )

# recode gender
table(dat_merged$female)

# recode income (originally coded during weighting process. Got from Lucid. Different from our question in the survey)
dat_merged <- dat_merged %>% 
  mutate(
    income_cat = case_when(
      inc5_cat == 1 ~ "<30",
      inc5_cat == 2 ~ "30-59",
      inc5_cat == 3 ~ "60-99",
      inc5_cat == 4 ~ "100-149",
      inc5_cat == 5 ~ "150+",
      TRUE ~ NA_character_
    ) %>% 
      factor(levels = c("<30", "30-59", "60-99", "100-149", "150+"))
  )

# Create dummy variables
dat_merged <- dat_merged %>% 
  mutate(
    income_30_less   = ifelse(income_cat == "<30", 1, 0),
    income_30_59     = ifelse(income_cat == "30-59", 1, 0),
    income_60_99     = ifelse(income_cat == "60-99", 1, 0),
    income_100_149   = ifelse(income_cat == "100-149", 1, 0),
    income_150_plus  = ifelse(income_cat == "150+", 1, 0)
  )
  
# recode veteran (Q84)
dat_merged <- dat_merged %>% 
  mutate(veteran = case_when(
    Q84 == 1 ~ 0,
    Q84 == 2 ~ 1,
    Q84 == 3 ~ 0,
    TRUE ~ NA_real_
  ))

# recode active duty (Q84)
dat_merged <- dat_merged %>% 
  mutate(active_milit = case_when(
    Q84 == 1 ~ 1,
    Q84 == 2 ~ 0,
    Q84 == 3 ~ 0,
    TRUE ~ NA_real_
  ))

# recode law_enforcement (Q85)
dat_merged <- dat_merged %>% 
  mutate(law_enforcement = case_when(
    Q85 == 1 ~ 1,
    Q85 == 2 ~ 1,
    Q85 == 3 ~ 0,
    TRUE ~ NA_real_
  ))

# recode days carried outside the home
dat_merged <- dat_merged %>% 
  mutate(days_carried = Q81 - 1)
# we'll definitely want to test these removing the extremes as well
# remove options for 30 days and 0 days
dat_merged <- dat_merged %>% 
  mutate(days_carried_no_high = ifelse(days_carried < 30, days_carried, NA),
         days_carried_no_extremes = ifelse(days_carried > 0 & days_carried < 30, days_carried, NA))

# oh also a dichotomous variable for whether you carried a gun in the last 30 days or not
dat_merged <- dat_merged %>% 
  mutate(carried_past_thirty_binary = case_when(
    days_carried == 0 ~ 0,
    days_carried > 0 ~ 1,
    TRUE ~ NA_real_
    ))

# make days_carried variable with 0 when they said they never carried in previous question
dat_merged <- dat_merged %>% 
  mutate(days_carried_include_nevers = case_when(
    !is.na(days_carried) ~ days_carried,
    Q80 == 1 ~ 0,
    TRUE ~ NA_real_)
  )

# code new carry frequency variable
dat_merged <- dat_merged %>%
  mutate(
    carry_freq = case_when(
      days_carried_include_nevers == 0 ~ 0,
      days_carried_include_nevers >= 1 & days_carried_include_nevers <= 14 ~ 1,
      days_carried_include_nevers >= 15 ~ 2,
      TRUE ~ NA_real_
    )
  )

table(dat_merged$carry_freq, useNA = "ifany")

dat_merged <- dat_merged %>%
  mutate(
    carry_freq_ord = factor(
      carry_freq,
      levels = c(0, 1, 2),
      labels = c("0 Days", "1-14 Days", "15-30 Days"),
      ordered = TRUE
    )
  )

table(dat_merged$carry_freq_ord, useNA = "ifany")


# make categorical carry_freq
dat_merged <- dat_merged %>%
  mutate(
    carry_freq_cat = factor(
      as.character(carry_freq_ord),
      levels = c("0 Days", "1-14 Days", "15-30 Days")  # specify order
    )
  )
str(dat_merged$carry_freq_cat)
table(dat_merged$carry_freq_cat)

# plan to purchase
sum(is.na(dat_merged$Q82))
dat_merged <- dat_merged %>% 
  mutate(purchase_plan = case_when(
    Q82 == 1 ~ 1,
    Q82 == 2 ~ 0,
    TRUE ~ NA_real_
  ))

# whether they have guns at home
dat_merged <- dat_merged %>% 
  mutate(gun_home = case_when(
    Q78 == 1 ~ 1,
    Q78 == 2 ~ 0,
    Q78 == 3 ~ 0,
    TRUE ~ NA_real_
  ))

dat_merged <- dat_merged %>% 
  mutate(gun_own = case_when(
    Q79 == 1 ~ 1,
    Q78 == 2 ~ 0, # including these so that if they said no guns in their home, it also means they don't own a gun. otherwise this variable would be *among people with guns in their home, do they own it*
    Q78 == 3 ~ 0,
    Q79 == 2 ~ 0,
    Q79 == 3 ~ 0,
    TRUE ~ NA_real_
  ))

# Fill in carry type variables from Q80 only where they are currently missing
# (so we don't overwrite manually coded values from open-ended responses)
dat_merged <- dat_merged %>%
  mutate(
    concealed_carry = case_when(
      is.na(concealed_carry) & Q80 == 2 ~ 1,   # Concealed carry
      is.na(concealed_carry) & Q80 %in% c(1, 3) ~ 0,
      TRUE ~ concealed_carry  # Keep existing manual codes
    ),
    open_carry = case_when(
      is.na(open_carry) & Q80 == 3 ~ 1,        # Open carry
      is.na(open_carry) & Q80 %in% c(1, 2) ~ 0,
      TRUE ~ open_carry
    ),
    no_carry = case_when(
      is.na(no_carry) & Q80 == 1 ~ 1,          # Does not carry
      is.na(no_carry) & Q80 %in% c(2, 3) ~ 0,
      TRUE ~ no_carry
    )
  )

# Create a collapsed binary carry variable
# 1 = carries (either concealed or open), 0 = has guns but does not carry outside home
dat_merged <- dat_merged %>%
  mutate(
    carry = case_when(
      open_carry == 1 | concealed_carry == 1 ~ 1,
      no_carry == 1 ~ 0, # has guns (not NA) but doesn't carry outside the home
      TRUE ~ NA_real_
    )
  )

# Create mutually exclusive carry type categorical
dat_merged <- dat_merged %>%
  mutate(
    carry_type_cat = case_when(
      open_carry == 1 ~ "Open carry",            # Highest priority
      concealed_carry == 1 ~ "Concealed carry",  # Second priority
      no_carry == 1 ~ "No carry",                # Lowest priority
      TRUE ~ NA_character_                       # Shouldn't happen, but protects integrity
    ) %>%
      factor(
        levels = c("No carry", "Concealed carry", "Open carry")
      )
  )

table(dat_merged$open_carry)
sum(is.na(dat_merged$open_carry))
table(dat_merged$concealed_carry)
sum(is.na(dat_merged$concealed_carry))
table(dat_merged$no_carry)
sum(is.na(dat_merged$no_carry))
nrow(dat_merged)

####  Create mutually exclusive carry type categorical ####
# Clean Q83 multiple-select: replace NAs with 0
dat_merged <- dat_merged %>%
  mutate(across(c(Q83_1, Q83_2, Q83_3, Q83_4, Q83_5),
                ~ tidyr::replace_na(., 0)))

# Create mutually exclusive firearm categories
dat_merged <- dat_merged %>%
  mutate(
    handgun = as.integer(Q83_1 == 1 | Q83_2 == 1),
    longgun = as.integer(Q83_3 == 1 | Q83_4 == 1 | Q83_5 == 1),
    gun_type_cat = case_when(
      handgun == 1 & longgun == 0 ~ "Handgun only",
      handgun == 0 & longgun == 1 ~ "Long gun only",
      handgun == 1 & longgun == 1 ~ "Mixed",
      TRUE                       ~ "None"
    )
  )

# Coerce to factor
dat_merged <- dat_merged %>%
  mutate(
    gun_type_cat = factor(
      gun_type_cat,
      levels = c("None",
                 "Handgun only",
                 "Long gun only",
                 "Mixed")
    )
  )

```

```{r Make variables for state permisiveness | Permit & Training Requirements}

# state permit and training requirements gathered from Eveytown website: https://maps.everytownresearch.org/navigator/states.html?dataset=permitting_process&states=NY#q-gunmath_concealed_permit_required

#### Permit Requirement ####

state_permissive <- read_csv("data/Everytown_CC_permissive states.csv")

state_permissive %>% 
  pull(state) %>% 
  n_distinct()

dat_merged %>% 
  pull(state) %>% 
  n_distinct()

# # which states are in dat_merged but not in state_permissive
# setdiff(unique(dat_merged$state), unique(state_permissive$state))
# # puerto rico is in dat_merged. Probably just have to put those as NA or cut from dataset

# dat_merged %>% 
#   filter(state == "PR") %>% 
#   nrow()
# # ah. okay it's one person. Sorry, bye puerto rico
# # putting code up front to remove people from PR
# cut PR in manual zipcode coding


table(state_permissive$response)
table(is.na(state_permissive$response))

# see states that don't require permits for CC
state_permissive %>% 
  filter(response == "No") %>% 
  pull(state)

# create vector directly from the table
concealed_permissive_states <- state_permissive %>%
  filter(response == "No") %>%
  pull(state)

# use it in the dataset
dat_merged <- dat_merged %>%
  mutate(cc_no_permit = if_else(
    state %in% concealed_permissive_states, 1, 0
  ))



dat_merged %>% 
  filter(cc_no_permit == 1) %>% 
  nrow()
dat_merged %>% 
  filter(cc_no_permit == 0) %>% 
  nrow()
# pretty balanced in terms of rows in permissive vs. permit states

#### Training Requirement ####

# get data for states that require training for cc

state_training <- read_csv("data/Everytown_CC_training.csv")

state_training %>% 
  pull(state) %>% 
  n_distinct()

dat_merged %>% 
  pull(state) %>% 
  n_distinct()

table(state_training$response)
table(is.na(state_training$response))

# see states that don't require training for CC
state_training %>% 
  filter(response == "No") %>% 
  pull(state)

# create vector directly from the table
concealed_permissive_states <- state_training %>%
  filter(response == "No") %>%
  pull(state)

# use it in the dataset
dat_merged <- dat_merged %>%
  mutate(cc_no_training = if_else(
    state %in% concealed_permissive_states, 1, 0
  ))



table(dat_merged$cc_no_permit, dat_merged$cc_no_training)


# make binary for permissive vs. restrictive
dat_merged <- dat_merged %>%
  mutate(state_permissiveness_01 = case_when(
    cc_no_permit == 0 & cc_no_training == 0 ~ 0,
    cc_no_permit == 1 & cc_no_training == 1 ~ 1,
    TRUE ~ NA_real_
  ))

# make continuous/ordinal permissiveness variable
dat_merged <- dat_merged %>%
  mutate(state_permissiveness_con = case_when(
    cc_no_permit == 0 & cc_no_training == 0 ~ 0,
    cc_no_permit == 0 & cc_no_training == 1 ~ .5,
    cc_no_permit == 1 & cc_no_training == 1 ~ 1,
    TRUE ~ NA_real_
  ))

```

```{r Identify states with low N}
# Note: resolve_state_name() is defined in Utility functions chunk

# Desired order for categories
lvl <- c("A lot (100+)",
         "Some (50–99)",
         "Few (20–49)",
         "Very few (<20)")

# Counts & categories
counts_all <- dat_merged %>%
  filter(!is.na(state)) %>%
  mutate(state_full = resolve_state_name(state)) %>%
  filter(!is.na(state_full)) %>%
  count(state_full, name = "n_all") %>%
  mutate(category_all = case_when(
    n_all < 20  ~ "Very few (<20)",
    n_all < 50  ~ "Few (20–49)",
    n_all < 100 ~ "Some (50–99)",
    TRUE        ~ "A lot (100+)"
  )) %>%
  mutate(category_all = factor(category_all, levels = lvl))

counts_gun <- dat_merged %>%
  filter(!is.na(state), Q79 == 1) %>%
  mutate(state_full = resolve_state_name(state)) %>%
  filter(!is.na(state_full)) %>%
  count(state_full, name = "n_gun") %>%
  mutate(category_gun = case_when(
    n_gun < 20  ~ "Very few (<20)",
    n_gun < 50  ~ "Few (20–49)",
    n_gun < 100 ~ "Some (50–99)",
    TRUE        ~ "A lot (100+)"
  )) %>%
  mutate(category_gun = factor(category_gun, levels = lvl))

# Add abbrevs for usmap (needed for later map plotting)
counts_all <- counts_all %>%
  mutate(state = state.abb[match(state_full, state.name)])
counts_gun <- counts_gun %>%
  mutate(state = state.abb[match(state_full, state.name)])

# see states with the smallest N
# states with smallest total sample size
counts_all %>%
  arrange(n_all) %>%
  dplyr::select(state_full, n_all, category_all) %>%
  slice_head(n = 10)

# states with smallest gun-owner sample size
counts_gun %>%
  arrange(n_gun) %>%
  dplyr::select(state_full, n_gun, category_gun) %>%
  slice_head(n = 20)

# see what we'd be losing if we dropped the very few states
# States to drop
drop_states <- counts_gun %>%
  filter(category_gun == "Very few (<20)") %>%
  pull(state_full)

# convert full names -> abbreviations
drop_states_abbr <- state.abb[match(drop_states, state.name)]
drop_states_abbr

# before dropping
table_before_permit <- dat_merged %>%
  distinct(state, cc_no_permit) %>%
  count(cc_no_permit, name = "states_before")

table_before_training <- dat_merged %>%
  distinct(state, cc_no_training) %>%
  count(cc_no_training, name = "states_before")

# after dropping "very few" states
table_after_permit <- dat_merged %>%
  filter(!state %in% drop_states_abbr) %>%
  distinct(state, cc_no_permit) %>%
  count(cc_no_permit, name = "states_after")

table_after_training <- dat_merged %>%
  filter(!state %in% drop_states_abbr) %>%
  distinct(state, cc_no_training) %>%
  count(cc_no_training, name = "states_after")

# merge before/after for each variable
full_join(table_before_permit, table_after_permit, by = "cc_no_permit") %>%
  arrange(cc_no_permit)

full_join(table_before_training, table_after_training, by = "cc_no_training") %>%
  arrange(cc_no_training)
```

```{r Display missingness of each relevant variable}
# Define all variables of interest
controls <- c(
  "age1829", "age3044", "age4564", "female", "hisp", "other_race", "black",
  "college", "gop", "ind", "notreligious", "otherreligion", "ideology",
  "veteran", "active_milit", "law_enforcement", "worrycrime", "victimcrime",
  "income_30_less", "income_60_99", "income_100_149", "income_150_plus", "econ"
)

state_vars <- c(
  "state_violent_rate_2024_z", "state_pct_poverty_2024_z",
  "state_pct_black_2020_z", "state_trump_perc_2024_z", "state_gun_own_2022_z"
)

predictors <- c("state", "cc_no_training", "cc_no_permit")
gun_owner_dvs <- c("carry", "open_carry", "carry_freq_ord")
all_dvs <- c("gun_own", "gun_home", "purchase_plan", "gun_type_cat")

all_vars <- unique(c(predictors, controls, state_vars, gun_owner_dvs, all_dvs))

# Compute NA counts and percentages
na_summary <- dat_merged %>%
  summarise(across(all_of(all_vars),
                   list(
                     missing_n = ~sum(is.na(.)),
                     missing_pct = ~mean(is.na(.)) * 100
                   ),
                   .names = "{.col}_{.fn}")) %>%
  pivot_longer(everything(),
               names_to = c("variable", ".value"),
               names_pattern = "(.*)_(missing_.*)") %>%
  arrange(desc(missing_pct))

# Display
print(na_summary, n = nrow(na_summary))
```

```{r Recode Demographics for Descriptives Tables}

dat_merged <- dat_merged %>% 
  mutate(gender_char = case_when(
    female == 0 ~ "Male",
    female == 1 ~ "Female",
    TRUE ~ NA_character_
  ))

dat_merged <- dat_merged %>% 
  mutate(race_char = case_when(
    white == 1 ~ "White",
    black == 1 ~ "Black",
    hisp == 1 ~ "Hispanic",
    other_race == 1 ~ "Other",
    TRUE ~ NA_character_
  ))

dat_merged <- dat_merged %>% 
  mutate(age_group_char = case_when(
    age1829 == 1 ~ "18-29",
    age3044 == 1 ~ "30-44",
    age4564 == 1 ~ "45-64",
    age65p == 1 ~ "65+",
    TRUE ~ NA_character_
  ))

dat_merged <- dat_merged %>% 
  mutate(religion_char = case_when(
    christian == 1 ~ "Christian (any denomination)",
    otherreligion == 1 ~ "Other Religion",
    notreligious == 1 ~ "Non-Religious",
    TRUE ~ NA_character_
  ))

dat_merged <- dat_merged %>% 
  mutate(college_char = case_when(
    college == 1 ~ "College-Educated",
    college == 0 ~ "Non-College",
    TRUE ~ NA_character_
  ))

dat_merged <- dat_merged %>% 
  mutate(party_char = case_when(
    gop == 1 ~ "Republican",
    ind == 1 ~ "Independent",
    dem == 1 ~ "Democrat",
    TRUE ~ NA_character_
  ))

# Income - already coded, but just called "_cat"
dat_merged <- dat_merged %>%
  mutate(
    income_char = factor(
      income_cat,
      levels = c("<30", "30-59", "60-99", "100-149", "150+")
    )
  )

# Income - already coded, but just called "_cat"
dat_merged <- dat_merged %>%
  mutate(
    income_char = case_when(
      income_cat == "<30" ~ "< $30,000",
      income_cat == "30-59" ~ "$30,000-59,000",
      income_cat == "60-99" ~ "$60,000-99,000",
      income_cat == "100-149" ~ "$100,000-149,000",
      income_cat == "150+" ~ "$150,000 +",
    TRUE ~ NA_character_
    ))
      

# veteran + active_milit + law_enforcement
dat_merged <- dat_merged %>% 
  mutate(veteran_char = case_when(
    veteran == 0 ~ "No",
    veteran == 1 ~ "Yes",
    TRUE ~ NA_character_
  ),
  law_enforcement_char = case_when(
    law_enforcement == 0 ~ "No",
    law_enforcement == 1 ~ "Yes",
    TRUE ~ NA_character_
  ),
  active_milit_char = case_when(
    active_milit == 0 ~ "No",
    active_milit == 1 ~ "Yes",
    TRUE ~ NA_character_
  ),)


dat_merged <- dat_merged %>% 
  mutate(permit_req_char = case_when(
    cc_no_permit == 0 ~ "Permit Required",
    cc_no_permit == 1 ~ "No Permit Required",
    TRUE ~ NA_character_
  ))

dat_merged <- dat_merged %>% 
  mutate(training_req_char = case_when(
    cc_no_training == 0 ~ "Training Required",
    cc_no_training == 1 ~ "No Training Required",
    TRUE ~ NA_character_
  ))


#### Order the levels of character demographic variables ####

dat_merged <- dat_merged %>%
  mutate(
    gender_char = factor(gender_char,
      levels = c("Male", "Female")
    ),
    
    race_char = factor(race_char,
      levels = c("White", "Black", "Hispanic", "Other")
    ),
    
    age_group_char = factor(age_group_char,
      levels = c("18-29", "30-44", "45-64", "65+")
    ),
    
    religion_char = factor(religion_char,
      levels = c("Christian (any denomination)",
                 "Other Religion",
                 "Non-Religious")
    ),
    
    college_char = factor(college_char,
      levels = c("Non-College", "College-Educated")
    ),
    
    income_char = factor(income_char,
      levels = c(
        "< $30,000",
        "$30,000-59,000",
        "$60,000-99,000",
        "$100,000-149,000",
        "$150,000 +"
      )
    ),
    
    party_char = factor(party_char,
      levels = c("Democrat", "Independent", "Republican")
    ),
    
    veteran_char = factor(veteran_char,
      levels = c("No", "Yes")
    ),
    
    law_enforcement_char = factor(law_enforcement_char,
      levels = c("No", "Yes")
    ),
    
    active_milit_char = factor(active_milit_char,
      levels = c("No", "Yes")
    ),
    
    permit_req_char = factor(permit_req_char,
      levels = c("Permit Required", "No Permit Required")
    ),
    
    training_req_char = factor(training_req_char,
      levels = c("Training Required", "No Training Required")
    )
  )
```

Relevant controls:
age1829 + age3044 + age65p + female + hisp + black + other_race + college + income_30_59 + income_60_99 + income_100_149 + income_150_plus + gop + ind + notreligious + otherreligion + ideology + income + veteran + active_milit + law_enforcement + victimcrime + worrycrime + econ

reference categories are the largest group of any demographic:
reference race = white
reference gender = male
reference religion = christian
reference pid = dem
reference age = age4564
reference income = income_30_less

Making single master datasets that can be used in each DV universe. We make a gun-owners dataset, and all dataset, and a no small N states gun-owners dataset and a no small n states all dataset. Each uses list-wise deletion for rows missing any key or control variables.

*Primary DVs for gun-owners:*
- carry
- open_carry
- carry_freq_cat
- carry_type_cat

*Primary DVs for all:*
- gun_own
- gun_home
- purchase_plan
- gun_type_cat

```{r Make processed datasets for analysis}

controls <- c(
  "age1829", "age3044", "age65p", "female", "hisp", "other_race", "black",
  "college", "gop", "ind", "notreligious", "otherreligion", "ideology",
  "veteran", "active_milit", "law_enforcement", "victimcrime", "worrycrime",
  "income_30_less", "income_60_99", "income_100_149", "income_150_plus",
  "econ", "state_violent_rate_2024_z", "state_pct_poverty_2024_z", 
  "state_pct_black_2020_z", "state_trump_perc_2024_z", "state_gun_own_2022_z"
)

# key DVs
key_vars_all   <- c("state", "cc_no_training", "cc_no_permit", controls)
key_vars_owners <- c("state", "cc_no_training", "cc_no_permit", controls)

# --- Gun-owner DVs ---
owner_dvs <- c("carry", "open_carry", "carry_freq_ord")

# --- All-sample DVs ---
all_dvs <- c("gun_own", "gun_home", "purchase_plan", "gun_type_cat")

#### Function: listwise delete on controls (optionally on DVs) ####
make_processed <- function(data,
                           dvs,
                           key_vars,
                           drop_dvs = TRUE) {
  # Base vars we always require to be non-missing
  base_vars <- key_vars
  
  # Optionally also require DVs to be non-missing
  all_vars <- if (drop_dvs) {
    unique(c(base_vars, dvs))
  } else {
    base_vars
  }
  
  data %>%
    dplyr::filter(
      dplyr::if_all(
        dplyr::all_of(all_vars),
        ~ !is.na(.x)
      )
    )
}

#### Make each dataset ####

# All respondents for all-sample DVs
dat_all_processed <- make_processed(
  data     = dat_merged,
  dvs      = all_dvs,
  key_vars = key_vars_all,
  drop_dvs = TRUE     # here we *do* want listwise deletion incl. DVs
)

# Gun owners only, for owner DVs
dat_gunowners_processed <- dat_merged %>%
  dplyr::filter(gun_own == 1) %>%
  make_processed(
    dvs      = owner_dvs,
    key_vars = key_vars_owners,
    drop_dvs = TRUE     # keep requiring non-missing owner DVs
  )

# Non-gun-owners: only clean on controls, ignore owner DVs (which are NA)
dat_NONgunowners_processed <- dat_merged %>%
  dplyr::filter(gun_own == 0) %>%
  make_processed(
    dvs      = owner_dvs,
    key_vars = key_vars_owners,
    drop_dvs = FALSE    # do *not* drop on owner DVs
  )

#### Exclude small-N states ####

dat_all_processed_nomin <- dat_all_processed %>%
  dplyr::filter(!(state %in% drop_states_abbr))

dat_gunowners_processed_nomin <- dat_gunowners_processed %>%
  dplyr::filter(!(state %in% drop_states_abbr))

dat_NONgunowners_processed_nomin <- dat_NONgunowners_processed %>%
  dplyr::filter(!(state %in% drop_states_abbr))
```

dat_all_processed
dat_all_processed_nomin
dat_gunowners_processed
dat_gunowners_processed_nomin

```{r Make maps with final dataframe}
# order
lvl <- c("A lot (100+)",
         "Some (50–99)",
         "Few (20–49)",
         "Very few (<20)")

# Counts & categories
counts_all <- dat_all_processed %>%
  filter(!is.na(state)) %>%
  mutate(state_full = resolve_state_name(state)) %>%
  filter(!is.na(state_full)) %>%
  count(state_full, name = "n_all") %>%
  mutate(category_all = case_when(
    n_all < 20  ~ "Very few (<20)",
    n_all < 50  ~ "Few (20–49)",
    n_all < 100 ~ "Some (50–99)",
    TRUE        ~ "A lot (100+)"
  )) %>%
  mutate(category_all = factor(category_all, levels = lvl))

counts_gun <- dat_gunowners_processed %>%
  filter(!is.na(state), Q79 == 1) %>%
  mutate(state_full = resolve_state_name(state)) %>%
  filter(!is.na(state_full)) %>%
  count(state_full, name = "n_gun") %>%
  mutate(category_gun = case_when(
    n_gun < 20  ~ "Very few (<20)",
    n_gun < 50  ~ "Few (20–49)",
    n_gun < 100 ~ "Some (50–99)",
    TRUE        ~ "A lot (100+)"
  )) %>%
  mutate(category_gun = factor(category_gun, levels = lvl))

# Add abbrevs for usmap
counts_all <- counts_all %>%
  mutate(state = state.abb[match(state_full, state.name)])
counts_gun <- counts_gun %>%
  mutate(state = state.abb[match(state_full, state.name)])

# Palette
pal <- c(
  "Very few (<20)" = "#f2f0f7",
  "Few (20–49)"    = "#cbc9e2",
  "Some (50–99)"   = "#9e9ac8",
  "A lot (100+)"   = "#6a51a3"
)

# Plots
p_all <- plot_usmap(data = counts_all,
                    values = "category_all",
                    regions = "states") +
  scale_fill_manual(values = pal,
                    breaks = lvl,          # enforce order
                    drop = FALSE,          # keep all categories if unused
                    na.value = "grey90",
                    na.translate = FALSE,  # hide NA from legend
                    name = "Respondents per state") +
  labs(title = " ") +
  theme(legend.position = "bottom")

p_gun <- plot_usmap(data = counts_gun,
                    values = "category_gun",
                    regions = "states") +
  scale_fill_manual(values = pal,
                    breaks = lvl,
                    drop = FALSE,
                    na.value = "grey90",
                    na.translate = FALSE,
                    name = "Gun owners per state") +
  labs(title = " ") +
  theme(legend.position = "bottom")

p_all
p_gun

# Save
ggsave("output/figures/map_all_respondents.png", p_all, width = 8, height = 5, dpi = 300)
ggsave("output/figures/map_gun_owners.png",    p_gun, width = 8, height = 5, dpi = 300)

# see states with the smallest N
counts_all %>%
  arrange(n_all) %>%
  dplyr::select(state_full, n_all, category_all) %>%
  slice_head(n = 10)

```

```{r See missing data in original dat_merged dataset}

# Variables used in all models

vars_used <- c(
  # main predictor
  "state_permissiveness_con", "cc_no_permit", "cc_no_training",

  # individual-level controls
  "age1829", "age3044", "age65p", "female",
  "hisp", "other_race", "black", "college",
  "income_30_less",
  "income_30_59", "income_60_99",
  "income_100_149", "income_150_plus",
  "gop", "ind", "notreligious", "otherreligion",
  "ideology", "veteran", "active_milit",
  "law_enforcement", "worrycrime", "victimcrime",

  # state controls
  "state_violent_rate_2024_z",
  "state_pct_poverty_2024_z",
  "state_pct_black_2020_z",
  "state_trump_perc_2024_z",
  "state_gun_own_2022_z",

  # random intercept variable
  "state",
  
  # DVs
 "gun_home","gun_own","purchase_plan", "gun_type_cat"
)

# If you're running this for a specific DV, add it:
# vars_used <- c("purchase_plan", vars_used)

# Count NAs in these variables

na_summary <- dat_merged %>%
  dplyr::select(all_of(vars_used)) %>%
  summarise(across(everything(), ~ sum(is.na(.)), .names = "missing_{col}")) %>%
  pivot_longer(everything(),
               names_to   = "variable",
               values_to  = "n_missing") %>%
  mutate(variable = str_replace(variable, "^missing_", "")) %>%
  arrange(desc(n_missing))

print(na_summary, n = nrow(na_summary))



table(dat_merged$Q5)
sum(is.na(dat_merged$Q5))
sum(is.na(dat_merged$Q4))

# Build the table
missing_table <- na_summary %>%
  mutate(
    pct_missing = round(n_missing / nrow(dat_merged) * 100, 2),
    variable = case_match(variable,
      "veteran" ~ "Veteran status",
      "active_milit" ~ "Active military status",
      "law_enforcement" ~ "Law enforcement status",
      "victimcrime" ~ "Crime victimization",
      "worrycrime" ~ "Worry about crime",
      "gun_own" ~ "Gun ownership",
      "state_gun_own_2022_z" ~ "State gun-ownership rate (z)",
      "gun_home" ~ "Gun in home",
      "purchase_plan" ~ "Purchase intention",
      "state_permissiveness_con" ~ "State permissiveness (continuous)",
      "cc_no_permit" ~ "No permit required",
      "cc_no_training" ~ "No training required",
      "age1829" ~ "Age 18–29",
      "age3044" ~ "Age 30–44",
      "age65p" ~ "Age 65+",
      "female" ~ "Female",
      "hisp" ~ "Hispanic",
      "other_race" ~ "Other race",
      "black" ~ "Black",
      "college" ~ "College graduate",
      "income_30_less" ~ "Income <$30,000",
      "income_30_59" ~ "Income $30,000–$59,999",
      "income_60_99" ~ "Income $60,000–$99,999",
      "income_100_149" ~ "Income $100,000–$149,999",
      "income_150_plus" ~ "Income $150,000+",
      "gop" ~ "Republican",
      "ind" ~ "Independent",
      "notreligious" ~ "Not religious",
      "otherreligion" ~ "Other religion",
      "ideology" ~ "Ideology (conservatism)",
      "state_violent_rate_2024_z" ~ "State violent crime rate (z)",
      "state_pct_poverty_2024_z" ~ "State poverty rate (z)",
      "state_pct_black_2020_z" ~ "State % Black (z)",
      "state_trump_perc_2024_z" ~ "State Trump vote share (z)",
      "state" ~ "State",
      "gun_type_cat" ~ "Firearm type",
      .default = variable
    )
  ) %>%
  rename(
    Variable = variable,
    `N Missing` = n_missing,
    `% Missing` = pct_missing
  )

# Create flextable
ft <- flextable(missing_table) %>%
  set_header_labels(Variable = "Variable", `N Missing` = "N Missing", `% Missing` = "% Missing") %>%
  colformat_double(j = "% Missing", digits = 2) %>%
  fontsize(size = 10, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  autofit() %>%
  set_caption("Missing Data by Analytic Variable (N = 18,251)")

# Export to Word
doc <- read_docx() %>%
  body_add_flextable(ft)
print(doc, target = "output/missing_data_table.docx")

```

```{r Make table for missing data for original datasets}
# Variable labels
var_labels <- c(
  "cc_no_permit" = "No permit required",
  "age1829" = "Age 18\u201329", "age3044" = "Age 30\u201344", "age65p" = "Age 65+",
  "female" = "Female", "hisp" = "Hispanic", "other_race" = "Other race",
  "black" = "Black", "college" = "College graduate",
  "income_30_less" = "Income <$30,000", "income_30_59" = "Income $30,000\u2013$59,999",
  "income_60_99" = "Income $60,000\u2013$99,999", "income_100_149" = "Income $100,000\u2013$149,999",
  "income_150_plus" = "Income $150,000+",
  "gop" = "Republican", "ind" = "Independent",
  "notreligious" = "Not religious", "otherreligion" = "Other religion",
  "ideology" = "Ideology (conservatism)",
  "veteran" = "Veteran status", "active_milit" = "Active military status",
  "law_enforcement" = "Law enforcement status",
  "worrycrime" = "Worry about crime", "victimcrime" = "Crime victimization",
  "state_violent_rate_2024_z" = "State violent crime rate (z)",
  "state_pct_poverty_2024_z" = "State poverty rate (z)",
  "state_pct_black_2020_z" = "State % Black (z)",
  "state_trump_perc_2024_z" = "State Trump vote share (z)",
  "state_gun_own_2022_z" = "State gun-ownership rate (z)",
  "state" = "State",
  "gun_own" = "Gun ownership", "gun_home" = "Gun in home",
  "purchase_plan" = "Purchase intention", "gun_type_cat" = "Firearm type",
  "carry" = "Carry outside home", "open_carry" = "Open carry",
  "carry_freq_cat" = "Carry frequency", "carry_type_cat" = "Carry type"
)

shared_vars <- c(
  "cc_no_permit", "age1829", "age3044", "age65p", "female",
  "hisp", "other_race", "black", "college",
  "income_30_less", "income_30_59", "income_60_99",
  "income_100_149", "income_150_plus",
  "gop", "ind", "notreligious", "otherreligion",
  "ideology", "veteran", "active_milit",
  "law_enforcement", "worrycrime", "victimcrime",
  "state_violent_rate_2024_z", "state_pct_poverty_2024_z",
  "state_pct_black_2020_z", "state_trump_perc_2024_z",
  "state_gun_own_2022_z", "state"
)

dvs_all <- c("gun_own", "gun_home", "purchase_plan", "gun_type_cat")
dvs_go <- c("carry", "open_carry", "carry_freq_cat", "carry_type_cat")
all_vars <- c(dvs_all, dvs_go, shared_vars)

dat_go <- dat_merged %>% filter(gun_own == 1)
n_all <- nrow(dat_merged)
n_go <- nrow(dat_go)

# Count missing in full sample (only vars that exist)
vars_in_all <- intersect(all_vars, names(dat_merged))
vars_in_go <- intersect(all_vars, names(dat_go))

miss_all <- sapply(vars_in_all, function(v) sum(is.na(dat_merged[[v]])))
miss_go <- sapply(vars_in_go, function(v) sum(is.na(dat_go[[v]])))

# Build combined table
combined <- tibble(variable = all_vars) %>%
  mutate(
    Variable = ifelse(variable %in% names(var_labels), var_labels[variable], variable),
    n_miss_all = ifelse(variable %in% names(miss_all), miss_all[variable], NA_integer_),
    n_miss_go  = ifelse(variable %in% names(miss_go), miss_go[variable], NA_integer_),
    `Full Sample` = ifelse(
      is.na(n_miss_all), "\u2014",
      paste0(n_miss_all, " (", sprintf("%.1f", n_miss_all / n_all * 100), "%)")
    ),
    `Gun Owners` = ifelse(
      is.na(n_miss_go), "\u2014",
      paste0(n_miss_go, " (", sprintf("%.1f", n_miss_go / n_go * 100), "%)")
    )
  ) %>%
  dplyr::select(Variable, `Full Sample`, `Gun Owners`)

# Flextable
ft <- flextable(combined) %>%
  set_header_labels(
    Variable = "Variable",
    `Full Sample` = paste0("Full Sample\n(N = ", format(n_all, big.mark = ","), ")"),
    `Gun Owners` = paste0("Gun Owners\n(N = ", format(n_go, big.mark = ","), ")")
  ) %>%
  align(j = 2:3, align = "center", part = "all") %>%
  fontsize(size = 10, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  autofit() %>%
  set_caption("Missing Data by Analytic Variable (N Missing / % of Sample)")

doc <- read_docx() %>%
  body_add_flextable(ft)
print(doc, target = "output/missing_data_table.docx")
```

```{r See missing data in full processed dataset}
# Variables used in all models

vars_used <- c(
  # main predictor
  "state_permissiveness_con", "cc_no_permit", "cc_no_training",

  # individual-level controls
  "age1829", "age3044", "age65p", "female",
  "hisp", "other_race", "black", "college",
  "gop", "ind", "notreligious", "otherreligion",
  "ideology", "veteran", "active_milit",
  "law_enforcement", "worrycrime", "victimcrime",

  # state controls
  "state_violent_rate_2024_z",
  "state_pct_poverty_2024_z",
  "state_pct_black_2020_z",
  "state_trump_perc_2024_z",
  "state_gun_own_2022_z",

  # random intercept variable
  "state",
  
  # DVs
 "gun_home","gun_own","purchase_plan", "gun_type_cat"
)

# If running this for a specific DV, add it:
# vars_used <- c("purchase_plan", vars_used)

# Count NAs in these variables

na_summary <- dat_all_processed %>%
  dplyr::select(all_of(vars_used)) %>%
  summarise(across(everything(), ~ sum(is.na(.)), .names = "missing_{col}")) %>%
  pivot_longer(everything(),
               names_to   = "variable",
               values_to  = "n_missing") %>%
  mutate(variable = str_replace(variable, "^missing_", "")) %>%
  arrange(desc(n_missing))

print(na_summary, n = nrow(na_summary))
```

```{r See missing data in gun owners processed dataset}
# Variables used in all models

vars_used <- c(
  # main predictors
  "state_permissiveness_con", "cc_no_permit", "cc_no_training",

  # individual-level controls
  "age1829", "age3044", "age65p", "female",
  "hisp", "other_race", "black", "college",
  "gop", "ind", "notreligious", "otherreligion",
  "ideology", "veteran", "active_milit",
  "law_enforcement", "worrycrime", "victimcrime",

  # state controls
  "state_violent_rate_2024_z",
  "state_pct_poverty_2024_z",
  "state_pct_black_2020_z",
  "state_trump_perc_2024_z",
  "state_gun_own_2022_z",

  # random intercept variable
  "state",
  
  # DVs
 "carry","open_carry", "carry_freq_ord"
)

# If running this for a specific DV, add it:
# vars_used <- c("purchase_plan", vars_used)


# Count NAs in these variables
na_summary <- dat_gunowners_processed %>%
  dplyr::select(all_of(vars_used)) %>%
  summarise(across(everything(), ~ sum(is.na(.)), .names = "missing_{col}")) %>%
  pivot_longer(everything(),
               names_to   = "variable",
               values_to  = "n_missing") %>%
  mutate(variable = str_replace(variable, "^missing_", "")) %>%
  arrange(desc(n_missing))

print(na_summary, n = nrow(na_summary))
```

```{r Shared analysis specifications}
#### Main IV ####
main_iv <- "cc_no_permit"

#### Control variables ####
control_vars <- c(
  "age1829", "age3044", "age65p", "female",
  "hisp", "black", "other_race",
  "college",
  "income_30_59", "income_60_99",
  "income_100_149", "income_150_plus",
  "gop", "ind",
  "notreligious", "otherreligion",
  "ideology",
  "veteran", "active_milit", "law_enforcement",
  "worrycrime", "victimcrime",
  "econ"
)

# Variant with gun ownership as additional control
control_vars_gunown <- c("gun_own", control_vars)

#### State covariate sets ####
state_sets <- list(
  set1 = "state_violent_rate_2024_z",
  set2 = "state_pct_poverty_2024_z",
  set3 = "state_pct_black_2020_z",
  set4 = "state_trump_perc_2024_z",
  set5 = "state_gun_own_2022_z"
)

#### Model titles ####
model_titles <- c(
  "Base",
  "+ Ind Controls",
  "+ Violent crime rate (state)",
  "+ Poverty rate (state)",
  "+ % Black (state)",
  "+ Trump 2024 vote share (state)",
  "+ Gun-ownership rate (state)"
)

#### Variable labels ####
var_labels <- c(
  "(Intercept)"                 = "Intercept",
  "cc_no_permit"                = "No-permit state",
  "age1829"                     = "Age 18–29",
  "age3044"                     = "Age 30–44",
  "age65p"                      = "Age 65+",
  "female"                      = "Female",
  "hisp"                        = "Hispanic",
  "black"                       = "Black",
  "other_race"                  = "Other race",
  "college"                     = "College graduate",
  "income_30_59"                = "Income ($30–59,999)",
  "income_60_99"                = "Income ($60–99,999)",
  "income_100_149"              = "Income ($100–149,999)",
  "income_150_plus"             = "Income ($150,000+)",
  "gop"                         = "Republican",
  "ind"                         = "Independent",
  "notreligious"                = "Not religious",
  "otherreligion"               = "Other religion",
  "ideology"                    = "Ideology (conservatism)",
  "veteran"                     = "Veteran",
  "active_milit"                = "Active duty",
  "law_enforcement"             = "Law enforcement",
  "worrycrime"                  = "Worried about crime",
  "victimcrime"                 = "Crime victim",
  "econ"                        = "Financial worry",
  "state_violent_rate_2024_z"   = "Violent crime rate (z)",
  "state_pct_poverty_2024_z"    = "Poverty % (z)",
  "state_pct_black_2020_z"      = "% Black (z)",
  "state_trump_perc_2024_z"     = "Trump vote % (z)",
  "state_gun_own_2022_z"        = "Gun ownership % (z)"
)

# Variant with gun ownership label
var_labels_gunown <- c(
  var_labels[1:2],  # Intercept, cc_no_permit
  "gun_own" = "Gun owner",
  var_labels[3:length(var_labels)]
)

#### Variable ordering ####
var_order <- c(
  "cc_no_permit",
  "age1829", "age3044", "age65p", "female",
  "hisp", "black", "other_race",
  "college",
  "income_30_59", "income_60_99",
  "income_100_149", "income_150_plus",
  "gop", "ind",
  "notreligious", "otherreligion",
  "ideology",
  "veteran", "active_milit", "law_enforcement",
  "worrycrime", "victimcrime", "econ",
  "state_violent_rate_2024_z",
  "state_pct_poverty_2024_z",
  "state_pct_black_2020_z",
  "state_trump_perc_2024_z",
  "state_gun_own_2022_z"
)

# Variant with gun ownership in order
var_order_gunown <- c(
  "cc_no_permit", "gun_own",
  var_order[2:length(var_order)]
)
```

```{r Remove large objects}
rm(output)
rm(results_list)
rm(response)
rm(hud_key)
rm(hud_zipkey)
rm(expected_from_flags)
rm(p_all)
rm(p_gun)
rm(key)
rm(table_after_permit)
rm(table_after_training)
rm(table_before_permit)
rm(table_before_training)
rm(poverty_join)
rm(trump_join)
rm(gunown_join)
rm(crime_join)
rm(black_join)
rm(dat_black)
rm(dat_crime)
rm(dat_gun_ownership)
rm(dat_poverty)
rm(dat_trump)
rm(model1)
rm(m1)
rm(mods)
rm(m_unw)
rm(m_wt)
rm(ft)
rm(assembled)

```

***Run everything above here***

Specify multivariate regression models to test whether gun owners in permitless states are more likely to carry firearms outside the home, carry them concealed or open, and how frequently they carry outside the home, including clusters at the state level.

***Descriptives***

```{r Some descriptives for the report text}
nrow(dat) - nrow(dat_all_processed)

prop.table(table(dat_gunowners_processed$carry_type_cat, useNA = "ifany"))*100

table(dat_gunowners_processed$Q80)

dat_gunowners_processed %>% 
  dplyr::count(Q80, prop = TRUE, drop = FALSE)

prop.table(table(dat_gunowners_processed$Q80, useNA = "ifany"))*100

table(dat_gunowners_processed$Q80_4_TEXT)

dat_gunowners_processed %>%
  count(Q80) %>% 
  mutate(pct = round(n / sum(n) *100,1))

dat_orig <- read_csv("data/data_weighted_stata.csv")
dat_orig %>%
  mutate(Q80_orig = ifelse(!is.na(Q80_4_TEXT), 4, Q80)) %>%
  mutate(Q80_orig = as.numeric(Q80_orig)) %>% 
  count(Q80_orig) %>% 
  mutate(pct = n / sum(n) *100)
rm(dat_orig)

```

Figure for days carried
```{r Histogram | Days Carried (Including Nevers)}
ggplot(dat_gunowners_processed,
       aes(x = days_carried_include_nevers)) +
  geom_histogram(
    binwidth = 1,
    color = "white",
    fill = "steelblue"
  ) +
  scale_x_continuous(
    breaks = seq(0, 30, 5)   # no limits → keeps all observations
  ) +
  labs(
    x = "Days carried (0–30)",
    y = "Count of respondents"
  ) +
  theme_classic(base_size = 14, base_family = "Times New Roman") +
  theme(
    panel.grid = element_blank(),
    text = element_text(family = "Times New Roman")
  )

table(dat_gunowners_processed$days_carried_include_nevers, useNA = "ifany")
```

Demographics Mega-Table

```{r Demographics | Master Table}
#### Demographic variables ####
demos <- c(
  "gender_char","race_char","age_group_char","religion_char",
  "college_char","income_char","party_char",
  "veteran_char","law_enforcement_char","active_milit_char",
  "permit_req_char","training_req_char"
)

#### Human-readable titles ####
demo_titles <- c(
  gender_char          = "Gender",
  race_char            = "Race/ethnicity",
  age_group_char       = "Age group",
  religion_char        = "Religion",
  college_char         = "Education",
  income_char          = "Income",
  party_char           = "Party ID",
  veteran_char         = "Veteran status",
  law_enforcement_char = "Law enforcement status",
  active_milit_char    = "Active military status",
  permit_req_char      = "State concealed-carry permit requirement",
  training_req_char    = "State concealed-carry training requirement"
)

#### Function: build one demographic row-block ####
summarize_demo <- function(data, dm) {

  # must exist
  if (!all(c(dm, "wgtpi", "gun_own") %in% names(data))) return(NULL)

  df <- data %>%
    filter(!is.na(.data[[dm]]), !is.na(wgtpi)) %>%
    mutate(
      Level = .data[[dm]] |> factor()  # respects existing factor-level ordering if present
    )

  if (nrow(df) == 0) return(NULL)

  #### Weighted whole-sample % ####
  dsgn <- svydesign(ids = ~1, data = df, weights = ~wgtpi)
  tabw <- svytable(~ Level, dsgn)
  prop_w <- as.numeric(tabw) / sum(tabw)

  wt_df <- tibble(
    Level = factor(names(tabw), levels = levels(df$Level)),
    `% (weighted)` = sprintf("%.1f%%", 100 * prop_w)
  )

  #### Unweighted Ns (whole sample) ####
  Ns_total <- df %>%
    count(Level, name = "N (unwtd)")

  #### Unweighted gun-owner % and Ns ####
  owner_df <- dat_gunowners_processed %>%
    filter(!is.na(.data[[dm]])) %>%
    mutate(Level = factor(.data[[dm]], levels = levels(df$Level))) %>%
    group_by(Level) %>%
    summarise(
      `Gun owners N (unwtd)` = n(),
      `Gun owners % (unwtd)` =
        sprintf("%.1f%%", 100 * n() / nrow(dat_gunowners_processed)),
      .groups = "drop"
    )

  #### Combine all outputs ####
  out <- Ns_total %>%
    full_join(wt_df, by = "Level") %>%
    full_join(owner_df, by = "Level") %>%
    mutate(
      Demographic = unname(demo_titles[dm]) %||% dm
    ) %>%
    dplyr::select(
      Demographic, Level,
      `% (weighted)`, `N (unwtd)`,
      `Gun owners % (unwtd)`, `Gun owners N (unwtd)`
    )

  out
}

#### Build full table ####
all_blocks <- lapply(demos, \(dm) summarize_demo(dat_all_processed, dm))

big_tab <- bind_rows(all_blocks) %>%
  mutate(
    Demographic = factor(Demographic, levels = unname(demo_titles[demos]))
  ) %>%
  arrange(Demographic, Level)

#### Total row ####
total_n_all <- nrow(dat_all_processed)
total_n_gun <- nrow(dat_gunowners_processed)

total_row <- tibble(
  Demographic = "Total",
  Level = "",
  `% (weighted)` = "",
  `N (unwtd)` = total_n_all,
  `Gun owners % (unwtd)` = "",
  `Gun owners N (unwtd)` = total_n_gun
)

big_tab <- bind_rows(big_tab, total_row)

#### Word export ####
doc <- read_docx()
doc <- body_add_par(doc, "Demographic overview (whole sample weighted; gun ownership unweighted)", style = "heading 1")
doc <- body_add_par(
  doc,
  "Columns: weighted % of whole sample by level; unweighted N of whole sample; unweighted % gun owners within level; unweighted N gun owners.",
  style = "Normal"
)
doc <- body_add_par(doc, "")

# formatting
ft <- flextable(big_tab)
thin <- fp_border(color = "#999999", width = 0.75)
ft <- theme_booktabs(ft)
ft <- fontsize(ft, size = 9, part = "all")
ft <- align(ft, j = 1, align = "left", part = "all")
ft <- align(ft, j = 2, align = "left", part = "all")
ft <- align(ft, j = 3:6, align = "center", part = "all")
ft <- border_remove(ft)
ft <- hline_top(ft,    part = "header", border = thin)
ft <- hline_bottom(ft, part = "header", border = thin)
ft <- hline_bottom(ft, part = "body",   border = thin)
ft <- autofit(ft)
ft <- merge_v(ft, j = "Demographic")
ft <- valign(ft, j = "Demographic", valign = "top")

doc <- body_add_flextable(doc, ft)
print(doc, target = "output/Demographics_master_table.docx")
message("Wrote: Demographics_master_table.docx")
```

Descriptives Tables (Appendix E)

```{r Descriptives | Prep}

# demos to loop over
demos <- c(
  "gender_char", "race_char", "age_group_char",
  "religion_char", "college_char", "income_char",
  "party_char", "veteran_char", "law_enforcement_char",
  "active_milit_char", "permit_req_char", "training_req_char"
)

# return structure of demographic variables in dat_all_processed
purrr::walk(
  demos,
  ~ {
    cat("\n\n-----------------------------\n")
    cat("Variable:", .x, "\n")
    cat("-----------------------------\n")
    str(dat_all_processed[[.x]])
  }
)

# readable titles for each demo
demo_titles <- c(
  gender_char           = "Gender",
  race_char        = "Race/ethnicity",
  age_group_char   = "Age group",
  religion_char    = "Religion",
  college_char     = "Education",
  income_char      = "Income",
  party_char       = "Party ID",
  veteran_char          = "Veteran status",
  law_enforcement_char  = "Law enforcement status",
  active_milit_char     = "Active military status",
  permit_req_char = "State concealed-carry permit requirement",
  training_req_char= "State concealed-carry training requirement"
)

# Convenience: keep a consistent demo ordering
demo_order <- unique(unname(demo_titles[demos]))

# Unweighted descriptive table for any categorical outcome
summarize_unweighted_demo <- function(data, outcome, dm) {
  if (!all(c(dm, outcome) %in% names(data))) return(NULL)
  
  df <- data %>%
    dplyr::filter(
      !is.na(.data[[dm]]),
      !is.na(.data[[outcome]])
    ) %>%
    dplyr::mutate(
      grp = .data[[dm]],
      y   = .data[[outcome]]
    )
  
  if (nrow(df) == 0) return(NULL)
  
  # Count table
  tab  <- table(df$grp, df$y)
  prop <- prop.table(tab, margin = 1)
  
  # Convert to wide %
  prop_df <- as.data.frame(prop) %>%
    tidyr::pivot_wider(
      names_from  = Var2,
      values_from = Freq,
      names_prefix = "%_"
    ) %>%
    dplyr::mutate(
      dplyr::across(
        dplyr::starts_with("%_"),
        ~ sprintf("%.1f%%", 100 * .x)
      )
    ) %>%
    dplyr::rename(Level = Var1)
  
  # N per group
  Ns <- df %>%
    dplyr::count(Level = grp, name = "N (unwtd)")
  
  # Final table
  out <- dplyr::left_join(prop_df, Ns, by = "Level") %>%
    dplyr::mutate(
      Demographic = unname(demo_titles[dm]) %||% dm
    ) %>%
    dplyr::select(Demographic, Level, dplyr::everything())
  
  out
}

# Weighted descriptive table for any categorical outcome
summarize_weighted_demo <- function(data, outcome, dm) {
  if (!all(c(dm, outcome, "wgtpi") %in% names(data))) return(NULL)
  
  df <- data %>%
    dplyr::filter(
      !is.na(.data[[dm]]),
      !is.na(.data[[outcome]]),
      !is.na(wgtpi)
    ) %>%
    dplyr::mutate(
      grp = .data[[dm]],
      y   = .data[[outcome]]
    )
  
  if (nrow(df) == 0) return(NULL)
  
  dsgn <- svydesign(ids = ~1, weights = ~wgtpi, data = df)
  
  tab  <- svytable(~ grp + y, dsgn)
  prop <- prop.table(tab, margin = 1)
  
  prop_df <- as.data.frame(prop) %>%
    tidyr::pivot_wider(
      names_from  = y,
      values_from = Freq,
      names_prefix = "%_"
    ) %>%
    dplyr::mutate(
      dplyr::across(
        dplyr::starts_with("%_"),
        ~ sprintf("%.1f%%", 100 * .x)
      )
    ) %>%
    dplyr::rename(Level = grp)
  
  Ns <- df %>%
    dplyr::count(Level = grp, name = "N (unwtd)")
  
  out <- dplyr::left_join(prop_df, Ns, by = "Level") %>%
    dplyr::mutate(
      Demographic = unname(demo_titles[dm]) %||% dm
    ) %>%
    dplyr::select(Demographic, Level, dplyr::everything())
  
  out
}

# Shared flextable formatting
make_demo_flextable <- function(big_tab) {
  thin <- officer::fp_border(color = "#999999", width = 0.75)
  
  ft <- flextable::flextable(big_tab) %>%
    flextable::theme_booktabs() %>%
    flextable::fontsize(size = 9, part = "all") %>%
    flextable::align(j = 1:2, align = "left",  part = "all") %>%
    flextable::align(j = 3:ncol(big_tab), align = "center", part = "all") %>%
    flextable::border_remove() %>%
    flextable::hline_top(part = "header", border = thin) %>%
    flextable::hline_bottom(part = "header", border = thin) %>%
    flextable::hline_bottom(part = "body", border = thin) %>%
    flextable::autofit() %>%
    flextable::merge_v(j = "Demographic") %>%
    flextable::valign(j = "Demographic", valign = "top")
  
  ft
}
```

```{r Descriptives Table | Gun Ownership (weighted)}
outcome <- "gun_own"

all_tabs <- lapply(
  demos,
  function(dm) {
    summarize_weighted_demo(
      data    = dat_all_processed,
      outcome = outcome,
      dm      = dm
    )
  }
)

big_tab <- dplyr::bind_rows(all_tabs) %>%
  dplyr::mutate(
    Demographic = factor(Demographic, levels = demo_order)
  ) %>%
  dplyr::arrange(Demographic, Level)

# Total row
dsgn_total <- svydesign(
  ids     = ~1,
  data    = dat_all_processed %>% dplyr::filter(!is.na(.data[[outcome]])),
  weights = ~wgtpi
)

tab_total  <- svytable(~ dat_all_processed[[outcome]], dsgn_total)
prop_total <- prop.table(tab_total)

total_row <- tibble::tibble(
  Demographic = "Total sample",
  Level       = "",
  !!!setNames(
    sprintf("%.1f%%", 100 * as.numeric(prop_total)),
    paste0("%_", names(prop_total))
  ),
  `N (unwtd)` = sum(!is.na(dat_all_processed[[outcome]]))
)

big_tab <- dplyr::bind_rows(big_tab, total_row)

# Export
doc <- read_docx()
doc <- body_add_par(
  doc,
  "Weighted percentages of respondents who report personally owning firearms by demographic",
  style = "heading 1"
)
doc <- body_add_par(
  doc,
  "Percentages are weighted; Ns are unweighted.",
  style = "Normal"
)
doc <- body_add_par(doc, "")

ft <- make_demo_flextable(big_tab)
doc <- body_add_flextable(doc, ft)

print(doc, target = paste0("output/descriptives_", outcome, "_weighted.docx"))
message("Wrote: descriptives_", outcome, "_weighted.docx")
```

```{r Descriptives Table | Gun in Home (weighted)}
outcome <- "gun_home"

all_tabs <- lapply(
  demos,
  function(dm) {
    summarize_weighted_demo(
      data    = dat_all_processed,
      outcome = outcome,
      dm      = dm
    )
  }
)

big_tab <- dplyr::bind_rows(all_tabs) %>%
  dplyr::mutate(
    Demographic = factor(Demographic, levels = demo_order)
  ) %>%
  dplyr::arrange(Demographic, Level)

# Total row
dsgn_total <- svydesign(
  ids     = ~1,
  data    = dat_all_processed,
  weights = ~wgtpi
)

tab_total  <- svytable(~ dat_all_processed[[outcome]], dsgn_total)
prop_total <- prop.table(tab_total)

total_row <- tibble::tibble(
  Demographic = "Total sample",
  Level       = "",
  !!!setNames(
    sprintf("%.1f%%", 100 * as.numeric(prop_total)),
    paste0("%_", names(prop_total))
  ),
  `N (unwtd)` = sum(!is.na(dat_all_processed[[outcome]]))
)

big_tab <- dplyr::bind_rows(big_tab, total_row)

doc <- read_docx()
doc <- body_add_par(
  doc,
  "Weighted percentages of respondents who report having a firearm in their household by demographic",
  style = "heading 1"
)
doc <- body_add_par(
  doc,
  "Percentages are weighted; Ns are unweighted.",
  style = "Normal"
)
doc <- body_add_par(doc, "")

ft <- make_demo_flextable(big_tab)
doc <- body_add_flextable(doc, ft)

print(doc, target = paste0("output/descriptives_", outcome, "_weighted.docx"))
message("Wrote: descriptives_", outcome, "_weighted.docx")
```

```{r Descriptives Table | Purchase Plan (weighted)}
outcome <- "purchase_plan"

all_tabs <- lapply(
  demos,
  function(dm) {
    summarize_weighted_demo(
      data    = dat_all_processed,
      outcome = outcome,
      dm      = dm
    )
  }
)

big_tab <- dplyr::bind_rows(all_tabs) %>%
  dplyr::mutate(
    Demographic = factor(Demographic, levels = demo_order)
  ) %>%
  dplyr::arrange(Demographic, Level)

# Total row
dsgn_total <- svydesign(
  ids     = ~1,
  data    = dat_all_processed,
  weights = ~wgtpi
)

tab_total  <- svytable(~ dat_all_processed[[outcome]], dsgn_total)
prop_total <- prop.table(tab_total)

total_row <- tibble::tibble(
  Demographic = "Total sample",
  Level       = "",
  !!!setNames(
    sprintf("%.1f%%", 100 * as.numeric(prop_total)),
    paste0("%_", names(prop_total))
  ),
  `N (unwtd)` = sum(!is.na(dat_all_processed[[outcome]]))
)

big_tab <- dplyr::bind_rows(big_tab, total_row)

doc <- read_docx()
doc <- body_add_par(
  doc,
  "Weighted percentages of respondents who intend to purchase a firearm by demographic",
  style = "heading 1"
)
doc <- body_add_par(
  doc,
  "Percentages are weighted; Ns are unweighted.",
  style = "Normal"
)
doc <- body_add_par(doc, "")

ft <- make_demo_flextable(big_tab)
doc <- body_add_flextable(doc, ft)

print(doc, target = paste0("output/descriptives_", outcome, "_weighted.docx"))
message("Wrote: descriptives_", outcome, "_weighted.docx")
```

```{r Descriptives Table | Gun Type (cat; weighted)}

outcome <- "gun_type_cat"

all_tabs <- lapply(
  demos,
  function(dm) {
    summarize_weighted_demo(
      data    = dat_all_processed,
      outcome = outcome,
      dm      = dm
    )
  }
)

# Combine + order by demographic and level
big_tab <- dplyr::bind_rows(all_tabs) %>%
  dplyr::mutate(
    Demographic = factor(Demographic, levels = demo_order)
  ) %>%
  dplyr::arrange(Demographic, Level)

# Total row (weighted %)
dsgn_total <- svydesign(
  ids     = ~1,
  data    = dat_all_processed %>% dplyr::filter(!is.na(.data[[outcome]])),
  weights = ~wgtpi
)

tab_total  <- svytable(~ get(outcome), dsgn_total)
prop_total <- prop.table(tab_total)

total_row <- tibble::tibble(
  Demographic = "Total sample",
  Level       = "",
  !!!setNames(
    sprintf("%.1f%%", 100 * as.numeric(prop_total)),
    paste0("%_", names(prop_total))
  ),
  `N (unwtd)` = sum(!is.na(dat_all_processed[[outcome]]))
)

big_tab <- dplyr::bind_rows(big_tab, total_row)

# Export to Word
doc <- read_docx()
doc <- body_add_par(
  doc,
  "Weighted percentages of firearm type categories by demographic",
  style = "heading 1"
)
doc <- body_add_par(
  doc,
  "Percentages are weighted; Ns are unweighted.",
  style = "Normal"
)
doc <- body_add_par(doc, "")

ft <- make_demo_flextable(big_tab)

doc <- body_add_flextable(doc, ft)

print(doc, target = paste0("output/descriptives_", outcome, "_weighted.docx"))
message("Wrote: descriptives_", outcome, "_weighted.docx")
```

```{r Descriptives Table | Carry Frequency (cat; gun owners)}

outcome <- "carry_freq_cat"

all_tabs <- lapply(
  demos,
  function(dm) {
    summarize_unweighted_demo(
      data    = dat_gunowners_processed,
      outcome = outcome,
      dm      = dm
    )
  }
)

big_tab <- dplyr::bind_rows(all_tabs) %>%
  dplyr::mutate(
    Demographic = factor(Demographic, levels = demo_order)
  ) %>%
  dplyr::arrange(Demographic, Level)

tab_total  <- table(dat_gunowners_processed[[outcome]])
prop_total <- prop.table(tab_total)

total_row <- tibble::tibble(
  Demographic = "Total sample",
  Level       = "",
  !!!setNames(
    sprintf("%.1f%%", 100 * as.numeric(prop_total)),
    paste0("%_", names(prop_total))
  ),
  `N (unwtd)` = sum(!is.na(dat_gunowners_processed[[outcome]]))
)

big_tab <- dplyr::bind_rows(big_tab, total_row)

doc <- read_docx()
doc <- body_add_par(
  doc,
  "Unweighted percentages of gun owners’ carry frequency in the past 30 days by demographic",
  style = "heading 1"
)
doc <- body_add_par(
  doc,
  "Percentages and Ns are unweighted.",
  style = "Normal"
)
doc <- body_add_par(doc, "")

ft <- make_demo_flextable(big_tab)
doc <- body_add_flextable(doc, ft)

print(
  doc,
  target = paste0("output/descriptives_", outcome, "_unweighted_gun_owners.docx")
)
message(
  "Wrote: descriptives_",
  outcome,
  "_unweighted_gun_owners.docx"
)
```

```{r Descriptives Table | Carry Type (cat; gun owners)}

outcome <- "carry_type_cat"

all_tabs <- lapply(
  demos,
  function(dm) {
    summarize_unweighted_demo(
      data    = dat_gunowners_processed,
      outcome = outcome,
      dm      = dm
    )
  }
)

big_tab <- dplyr::bind_rows(all_tabs) %>%
  dplyr::mutate(
    Demographic = factor(Demographic, levels = demo_order)
  ) %>%
  dplyr::arrange(Demographic, Level)

tab_total  <- table(dat_gunowners_processed[[outcome]])
prop_total <- prop.table(tab_total)

total_row <- tibble::tibble(
  Demographic = "Total sample",
  Level       = "",
  !!!setNames(
    sprintf("%.1f%%", 100 * as.numeric(prop_total)),
    paste0("%_", names(prop_total))
  ),
  `N (unwtd)` = sum(!is.na(dat_gunowners_processed[[outcome]]))
)

big_tab <- dplyr::bind_rows(big_tab, total_row)

doc <- read_docx()
doc <- body_add_par(
  doc,
  "Unweighted percentages of gun owners’ carry type by demographic",
  style = "heading 1"
)
doc <- body_add_par(
  doc,
  "Percentages and Ns are unweighted.",
  style = "Normal"
)
doc <- body_add_par(doc, "")

ft <- make_demo_flextable(big_tab)
doc <- body_add_flextable(doc, ft)

print(
  doc,
  target = paste0("output/descriptives_", outcome, "_unweighted_gun_owners.docx")
)
message(
  "Wrote: descriptives_",
  outcome,
  "_unweighted_gun_owners.docx"
)
```

```{r Descriptives Table | Carry (concealed OR open; gun owners)}

outcome <- "carry"

all_tabs <- lapply(
  demos,
  function(dm) {
    summarize_unweighted_demo(
      data    = dat_gunowners_processed,
      outcome = outcome,
      dm      = dm
    )
  }
)

big_tab <- dplyr::bind_rows(all_tabs) %>%
  dplyr::mutate(
    Demographic = factor(Demographic, levels = demo_order)
  ) %>%
  dplyr::arrange(Demographic, Level)

tab_total  <- table(dat_gunowners_processed[[outcome]])
prop_total <- prop.table(tab_total)

total_row <- tibble::tibble(
  Demographic = "Total sample",
  Level       = "",
  `% (No)`    = sprintf("%.1f%%", 100 * prop_total["0"]),
  `% (Yes)`   = sprintf("%.1f%%", 100 * prop_total["1"]),
  `N (unwtd)` = sum(!is.na(dat_gunowners_processed[[outcome]]))
)

big_tab <- dplyr::bind_rows(big_tab, total_row)

doc <- read_docx()
doc <- body_add_par(
  doc,
  "Unweighted percentages of gun owners who report carrying firearms in public (concealed or open) by demographic",
  style = "heading 1"
)
doc <- body_add_par(
  doc,
  "Percentages and Ns are unweighted.",
  style = "Normal"
)
doc <- body_add_par(doc, "")

ft <- make_demo_flextable(big_tab)
doc <- body_add_flextable(doc, ft)

print(
  doc,
  target = paste0("output/descriptives_", outcome, "_unweighted_gun_owners.docx")
)
message(
  "Wrote: descriptives_",
  outcome,
  "_unweighted_gun_owners.docx"
)
```

```{r Descriptives Table | Concealed Carry (gun owners)}

outcome <- "concealed_carry"

all_tabs <- lapply(
  demos,
  function(dm) {
    summarize_unweighted_demo(
      data    = dat_gunowners_processed,
      outcome = outcome,
      dm      = dm
    )
  }
)

big_tab <- dplyr::bind_rows(all_tabs) %>%
  dplyr::mutate(
    Demographic = factor(Demographic, levels = demo_order)
  ) %>%
  dplyr::arrange(Demographic, Level)

tab_total  <- table(dat_gunowners_processed[[outcome]])
prop_total <- prop.table(tab_total)

total_row <- tibble::tibble(
  Demographic = "Total sample",
  Level       = "",
  `% (No)`    = sprintf("%.1f%%", 100 * prop_total["0"]),
  `% (Yes)`   = sprintf("%.1f%%", 100 * prop_total["1"]),
  `N (unwtd)` = sum(!is.na(dat_gunowners_processed[[outcome]]))
)

big_tab <- dplyr::bind_rows(big_tab, total_row)

doc <- read_docx()
doc <- body_add_par(
  doc,
  "Unweighted percentages of gun owners who report concealed carry of firearms in public by demographic",
  style = "heading 1"
)
doc <- body_add_par(
  doc,
  "Percentages and Ns are unweighted.",
  style = "Normal"
)
doc <- body_add_par(doc, "")

ft <- make_demo_flextable(big_tab)
doc <- body_add_flextable(doc, ft)

print(
  doc,
  target = paste0("output/descriptives_", outcome, "_unweighted_gun_owners.docx")
)
message(
  "Wrote: descriptives_",
  outcome,
  "_unweighted_gun_owners.docx"
)
```

```{r Descriptives Table | Open Carry (gun owners)}

outcome <- "open_carry"

all_tabs <- lapply(
  demos,
  function(dm) {
    summarize_unweighted_demo(
      data    = dat_gunowners_processed,
      outcome = outcome,
      dm      = dm
    )
  }
)

big_tab <- dplyr::bind_rows(all_tabs) %>%
  dplyr::mutate(
    Demographic = factor(Demographic, levels = demo_order)
  ) %>%
  dplyr::arrange(Demographic, Level)

tab_total  <- table(dat_gunowners_processed[[outcome]])
prop_total <- prop.table(tab_total)

total_row <- tibble::tibble(
  Demographic = "Total sample",
  Level       = "",
  `% (No)`    = sprintf("%.1f%%", 100 * prop_total["0"]),
  `% (Yes)`   = sprintf("%.1f%%", 100 * prop_total["1"]),
  `N (unwtd)` = sum(!is.na(dat_gunowners_processed[[outcome]]))
)

big_tab <- dplyr::bind_rows(big_tab, total_row)

doc <- read_docx()
doc <- body_add_par(
  doc,
  "Unweighted percentages of gun owners who report open carry of firearms in public by demographic",
  style = "heading 1"
)
doc <- body_add_par(
  doc,
  "Percentages and Ns are unweighted.",
  style = "Normal"
)
doc <- body_add_par(doc, "")

ft <- make_demo_flextable(big_tab)
doc <- body_add_flextable(doc, ft)

print(
  doc,
  target = paste0("output/descriptives_", outcome, "_unweighted_gun_owners.docx")
)
message(
  "Wrote: descriptives_",
  outcome,
  "_unweighted_gun_owners.docx"
)
```

***CLUSTERED SEs Word Tables of Main Models***
Appendix H, J, K, L, M, N, O

```{r Master Analyses Code | Functions and Specs}
# Note: control_vars, state_sets, model_titles, var_labels, and var_order
# are defined in the "Shared analysis specifications" chunk above

#### Build controls formula ####
controls_formula <- paste(control_vars, collapse = " + ")

#### Build model formulas (shared structure) ####
build_formulas <- function(dv) {
  f0 <- as.formula(glue("{dv} ~ {main_iv}"))
  f1 <- as.formula(glue("{dv} ~ {main_iv} + {controls_formula}"))
  f_states <- lapply(state_sets, function(sc) {
    as.formula(glue("{dv} ~ {main_iv} + {controls_formula} + {sc}"))
  })
  out <- c(list(f0, f1), f_states)
  names(out) <- model_titles
  out
}

#### Logistic models: survey design with state clustering ####
fit_survey_clustered <- function(fml, df, weight_var = NULL) {
  
  if (!is.null(weight_var)) {
    df <- df %>%
      filter(
        !is.na(.data[[weight_var]]),
        !is.na(state)
      )
  } else {
    df <- df %>%
      filter(!is.na(state))
  }
  
  if (!is.null(weight_var)) {
    w_fml <- as.formula(paste0("~", weight_var))
    design <- svydesign(
      ids     = ~state,
      weights = w_fml,
      data    = df
    )
  } else {
    design <- svydesign(
      ids     = ~state,
      weights = NULL,
      data    = df
    )
  }
  
  fit <- svyglm(
    formula = fml,
    design  = design,
    family  = quasibinomial()
  )
  
  broom::tidy(fit, conf.int = FALSE) %>%
    mutate(
      OR    = exp(estimate),
      SE    = std.error,
      p     = p.value,
      stars = sig_stars(p)
    ) %>%
    dplyr::select(term, OR, SE, p, stars)
}

#### Multinomial: clustered vcov helper ####
vcov_cluster_multinom <- function(model, cluster) {
  
  mf <- model$model
  if (is.null(mf)) {
    stop("Model must be fit with model = TRUE")
  }
  
  X <- model.matrix(terms(model), data = mf)
  fitted_probs <- fitted(model)
  y <- model.response(mf)
  
  y_levels <- levels(y)
  n <- nrow(X)
  k <- length(y_levels)
  
  Y_mat <- matrix(0, n, k)
  for (i in seq_len(k)) {
    Y_mat[, i] <- as.numeric(y == y_levels[i])
  }
  
  p <- ncol(X)
  k_minus_1 <- k - 1
  
  scores <- matrix(0, n, p * k_minus_1)
  for (i in seq_len(n)) {
    for (j in seq_len(k_minus_1)) {
      idx_start <- (j - 1) * p + 1
      idx_end   <- j * p
      resid <- Y_mat[i, j + 1] - fitted_probs[i, j + 1]
      scores[i, idx_start:idx_end] <- resid * X[i, ]
    }
  }
  
  clusters <- unique(cluster)
  n_clusters <- length(clusters)
  
  cluster_scores <- matrix(0, n_clusters, ncol(scores))
  for (c in seq_len(n_clusters)) {
    idx <- which(cluster == clusters[c])
    cluster_scores[c, ] <- colSums(scores[idx, , drop = FALSE])
  }
  
  meat <- crossprod(cluster_scores)
  vcov_model <- vcov(model)
  
  adj <- n_clusters / (n_clusters - 1) *
    (n - 1) / (n - p * k_minus_1)
  
  vcov_model %*% meat %*% vcov_model * adj
}

#### Multinomial: optional diagnostics for clustering ####
verify_clustered_ses <- function(model, cluster_var, data) {
  
  se_standard <- sqrt(diag(vcov(model)))
  vcov_robust <- vcov_cluster_multinom(model, cluster = data[[cluster_var]])
  se_clustered <- sqrt(diag(vcov_robust))
  
  se_comparison <- data.frame(
    Parameter    = names(se_standard),
    Standard_SE  = se_standard,
    Clustered_SE = se_clustered,
    Ratio        = se_clustered / se_standard
  )
  
  cat("\n==============================\n")
  cat("CLUSTERED SE DIAGNOSTICS\n")
  cat("==============================\n\n")
  cat("Mean ratio:", sprintf("%.3f", mean(se_comparison$Ratio)), "\n")
  cat("Min ratio:", sprintf("%.3f", min(se_comparison$Ratio)), "\n")
  cat("Max ratio:", sprintf("%.3f", max(se_comparison$Ratio)), "\n\n")
  
  list(
    se_comparison = se_comparison,
    vcov_robust   = vcov_robust
  )
}

#### Multinomial: fit function (clustered or unclustered) ####
fit_multinom_generic <- function(
  fml,
  data,
  cluster_var = "state",
  clustered   = TRUE,
  verify      = FALSE
) {
  
  data <- data %>%
    filter(!is.na(!!rlang::sym(cluster_var)))
  
  m <- multinom(fml, data = data, trace = FALSE, model = TRUE)
  
  if (clustered) {
    if (verify) {
      ver <- verify_clustered_ses(m, cluster_var, data)
      vc <- ver$vcov_robust
    } else {
      vc <- vcov_cluster_multinom(m, cluster = data[[cluster_var]])
    }
  } else {
    vc <- vcov(m)
  }
  
  coefs <- coef(m)
  if (!is.matrix(coefs)) {
    coefs <- matrix(coefs, nrow = 1)
    rownames(coefs) <- levels(m$y)[2]
  }
  
  outcome_levels  <- rownames(coefs)
  predictor_names <- colnames(coefs)
  n_pred          <- length(predictor_names)
  
  se_vec <- sqrt(diag(vc))
  
  results_list <- list()
  
  for (i in seq_len(nrow(coefs))) {
    outcome <- outcome_levels[i]
    idx_start <- (i - 1) * n_pred + 1
    idx_end   <- i * n_pred
    
    results_list[[outcome]] <- data.frame(
      y.level    = outcome,
      term       = predictor_names,
      estimate   = coefs[i, ],
      std.error  = se_vec[idx_start:idx_end],
      stringsAsFactors = FALSE
    ) %>%
      mutate(
        RRR   = exp(estimate),
        SE    = std.error,
        z     = estimate / std.error,
        p     = 2 * pnorm(abs(z), lower.tail = FALSE),
        stars = sig_stars(p)
      ) %>%
      dplyr::select(y.level, term, RRR, SE, stars)
  }
  
  bind_rows(results_list)
}

#### Format logistic results: one row per term, one column per model ####
format_results_wide_logit <- function(results_list) {
  
  processed <- purrr::imap(results_list, function(df, model_name) {
    df %>%
      mutate(
        Term     = var_labels[term] %||% term,
        or_value  = sprintf("%.2f\n(%.2f)", OR, SE),
        star_val  = stars
      ) %>%
      dplyr::select(term, Term, or_value, star_val) %>%
      rename(
        !!model_name           := or_value,
        !!paste0(model_name,
                 "_stars")     := star_val
      )
  })
  
  all_terms <- data.frame(
    term = c("(Intercept)", var_order),
    stringsAsFactors = FALSE
  ) %>%
    mutate(Term = var_labels[term] %||% term)
  
  result_df <- all_terms
  
  for (i in seq_along(processed)) {
    result_df <- result_df %>%
      full_join(
        processed[[i]] %>%
          dplyr::select(term, everything(), -Term),
        by = "term"
      )
  }
  
  result_df %>%
    mutate(
      order_index = ifelse(term == "(Intercept)", 0,
                           match(term, var_order))
    ) %>%
    filter(!is.na(order_index)) %>%
    arrange(order_index) %>%
    dplyr::select(-order_index, -term)
}

#### Format multinomial results: stacked by outcome, cc_no_permit first ####
format_results_wide_multinom <- function(results_list) {
  
  formatted <- imap(results_list, function(df, model_name) {
    df %>%
      mutate(
        Term     = var_labels[term] %||% term,
        rrr_disp  = sprintf("%.2f\n(%.2f)", RRR, SE),
        star_disp = stars
      ) %>%
      dplyr::select(y.level, term, Term, rrr_disp, star_disp) %>%
      rename(
        !!model_name           := rrr_disp,
        !!paste0(model_name,
                 "_stars")     := star_disp
      )
  })
  
  all_terms <- data.frame(
    term = c("(Intercept)", var_order),
    stringsAsFactors = FALSE
  ) %>%
    mutate(Term = var_labels[term] %||% term)
  
  all_outcomes <- unique(formatted[[1]]$y.level)
  wide_by_outcome <- list()
  
  for (outcome in all_outcomes) {
    
    outcome_data <- all_terms
    
    for (i in seq_along(formatted)) {
      model_data <- formatted[[i]] %>%
        filter(y.level == outcome) %>%
        dplyr::select(-y.level, -Term)
      
      outcome_data <- outcome_data %>%
        left_join(model_data, by = "term")
    }
    
    outcome_data <- outcome_data %>%
      mutate(
        is_cc = ifelse(term == "cc_no_permit", 0, 1),
        order_index = ifelse(
          term == "cc_no_permit",
          0,
          match(term, var_order)
        )
      ) %>%
      arrange(is_cc, order_index) %>%
      dplyr::select(-is_cc, -order_index) %>%
      mutate(y.level = outcome) %>%
      dplyr::select(y.level, Term, everything(), -term)
    
    wide_by_outcome[[outcome]] <- outcome_data
  }
  
  bind_rows(wide_by_outcome)
}

#### Word export for logistic models ####
run_logit_models_to_word <- function(
  dv,
  data,
  weight_var = NULL,
  file_name
) {
  
  formulas <- build_formulas(dv)
  
  weight_status <- if (!is.null(weight_var)) "Weighted" else "Unweighted"
  message("\n", paste(rep("=", 60), collapse = ""))
  message(weight_status, " analysis with state clustering for DV: ", dv)
  message(paste(rep("=", 60), collapse = ""), "\n")
  
  results_list <- imap(
    formulas,
    function(fml, nm) {
      message("Fitting model: ", nm)
      fit_survey_clustered(
        fml,
        df         = data,
        weight_var = weight_var
      )
    }
  )
  
  final_table <- format_results_wide_logit(results_list)
  ft <- flextable(final_table)
  
  # Remove header text for stars columns
  star_cols <- grep("_stars$", names(final_table))
  for (col_idx in star_cols) {
    ft <- set_header_labels(
      ft,
      values = setNames("", names(final_table)[col_idx])
    )
  }
  
    # ---- FIRST autofit for row height only ----
  ft <- autofit(ft)

  # ---- THEN apply column widths (autofit can no longer override them) ----

  ncols <- ncol(final_table)

  # 1) Term column
  ft <- width(ft, j = 1, width = 1.4)

  # 2) Default width for all other model OR columns
  ft <- width(ft, j = 2:ncols, width = 0.85)

  # 3) Odd-numbered columns (3,5,7,...) get narrow width
  #    these correspond to OR(SE) columns
  odd_cols <- seq(3, min(18, ncols), by = 2)
  ft <- width(ft, j = odd_cols, width = 0.35)

  # ---- Now alignment / fonts ----
  ft <- bold(ft, part = "header")
  ft <- align(ft, align = "center", part = "header")
  ft <- align(ft, j = 1, align = "left", part = "body")
  ft <- align(ft, j = -1, align = "center", part = "body")

  # Set font defaults
  ft <- font(ft, fontname = "Times New Roman", part = "all")
  ft <- fontsize(ft, size = 10, part = "all")
  
  doc <- read_docx()
  doc <- body_add_par(
    doc,
    glue("{weight_status} clustered logit models: {dv}"),
    style = "heading 1"
  )
  doc <- body_add_flextable(doc, ft)
  
  print(doc, target = file_name)
  message("✓ Wrote: ", file_name)
}

#### Word export for multinomial models ####
run_multinom_models_to_word <- function(
  dv,
  data,
  file_name,
  clustered = TRUE,
  verify_ses = FALSE
) {
  
  formulas <- build_formulas(dv)
  
  results_list <- imap(
    formulas,
    function(fml, nm) {
      message("Fitting: ", nm)
      fit_multinom_generic(
        fml,
        data       = data,
        clustered  = clustered,
        verify     = verify_ses
      )
    }
  )
  
  final_table <- format_results_wide_multinom(results_list)
  ft <- flextable(final_table)
  
  # Remove header text for stars columns
  star_cols <- grep("_stars$", names(final_table))
  for (col_idx in star_cols) {
    ft <- set_header_labels(
      ft,
      values = setNames("", names(final_table)[col_idx])
    )
  }
  
    # ---- Apply human-readable model titles to estimate columns ----
  est_cols <- which(!grepl("_stars$", names(final_table)) &
                     !names(final_table) %in% c("y.level", "Term"))

  # Sanity check: number of estimate columns should match model_titles
  stopifnot(length(est_cols) == length(model_titles))

  ft <- set_header_labels(
    ft,
    values = setNames(model_titles, names(final_table)[est_cols])
  )
  
  # ---- ONE autofit pass only ----
  ft <- autofit(ft)
  
  # ---- Explicit column widths (override autofit) ----
  ncols <- ncol(final_table)
  
  # Outcome label column
  ft <- width(ft, j = 1, width = 1.0)
  
  # Term column
  ft <- width(ft, j = 2, width = 1.4)
  
  # Default width for remaining columns
  if (ncols >= 3) {
    ft <- width(ft, j = 3:ncols, width = 0.85)
  }
  
  # Narrow estimate columns (3, 5, 7, ...)
  odd_cols <- seq(4, ncols, by = 2)
  ft <- width(ft, j = odd_cols, width = 0.35)
  
  # ---- Styling ----
  ft <- bold(ft, part = "header")
  ft <- align(ft, align = "center", part = "header")
  ft <- align(ft, j = 1:2, align = "left",  part = "body")
  ft <- align(ft, j = -(1:2), align = "center", part = "body")
  
  # Font defaults
  ft <- font(ft, fontname = "Times New Roman", part = "all")
  ft <- fontsize(ft, size = 10, part = "all")
  
  doc <- read_docx()
  se_type <- if (clustered) "State-clustered" else "Standard (unclustered)"
  
  doc <- body_add_par(
    doc,
    glue("Multinomial logit with {se_type} SEs: {dv}"),
    style = "heading 1"
  )
  doc <- body_add_par(
    doc,
    glue("Baseline category: {levels(data[[dv]])[1]}"),
    style = "Normal"
  )
  doc <- body_add_flextable(doc, ft)
  
  print(doc, target = file_name)
  message("✓ Wrote Word document: ", file_name)
}
```

```{r Master Analyses Code | Create Main Model Documents}
#### Usage: logistic models (weighted + unweighted) ####
run_logit_models_to_word(
  dv         = "gun_own",
  data       = dat_all_processed,
  weight_var = "wgtpi",
  file_name  = "output/clustered_models_gun_own_WEIGHTED.docx"
)

run_logit_models_to_word(
  dv         = "gun_own",
  data       = dat_all_processed,
  weight_var = NULL,
  file_name  = "output/clustered_models_gun_own_UNWEIGHTED.docx"
)

run_logit_models_to_word(
  dv         = "purchase_plan",
  data       = dat_NONgunowners_processed,
  weight_var = NULL,
  file_name  = "output/clustered_models_purchase_plan_UNWEIGHTED_non-owners.docx"
)

run_logit_models_to_word(
  dv         = "purchase_plan",
  data       = dat_gunowners_processed,
  weight_var = NULL,
  file_name  = "output/clustered_models_purchase_plan_UNWEIGHTED_owners.docx"
)

run_logit_models_to_word(
  dv         = "gun_home",
  data       = dat_all_processed,
  weight_var = "wgtpi",
  file_name  = "output/clustered_models_gun_home_WEIGHTED.docx"
)

run_logit_models_to_word(
  dv         = "gun_home",
  data       = dat_all_processed,
  weight_var = NULL,
  file_name  = "output/clustered_models_gun_home_UNWEIGHTED.docx"
)

#### Usage: logistic models for partial samples (unweighted) ####
run_logit_models_to_word(
  dv         = "carry",
  data       = dat_gunowners_processed,
  weight_var = NULL,
  file_name  = "output/clustered_models_carry_UNWEIGHTED.docx"
)

run_logit_models_to_word(
  dv         = "open_carry",
  data       = dat_gunowners_processed,
  weight_var = NULL,
  file_name  = "output/clustered_models_open_carry_UNWEIGHTED.docx"
)

run_logit_models_to_word(
  dv         = "carry",
  data       = dat_gunowners_processed_nomin,
  weight_var = NULL,
  file_name  = "output/clustered_models_carry_UNWEIGHTED_nomin.docx"
)

run_logit_models_to_word(
  dv         = "open_carry",
  data       = dat_gunowners_processed_nomin,
  weight_var = NULL,
  file_name  = "output/clustered_models_open_carry_UNWEIGHTED_nomin.docx"
)

#### Usage: multinomial models ####

## Gun type, all respondents (clustered)
run_multinom_models_to_word(
  dv         = "gun_type_cat",
  data       = dat_all_processed,
  file_name  = "output/models_gun_type_multinomial_CLUSTERED.docx",
  clustered  = TRUE,
  verify_ses = FALSE
)

## Gun type, non-gun-owners (clustered)
run_multinom_models_to_word(
  dv         = "gun_type_cat",
  data       = dat_NONgunowners_processed,
  file_name  = "output/models_gun_type_multinomial_CLUSTERED_non-owners.docx",
  clustered  = TRUE,
  verify_ses = FALSE
)

## Gun type, all respondents (unclustered)
run_multinom_models_to_word(
  dv         = "gun_type_cat",
  data       = dat_all_processed,
  file_name  = "output/models_gun_type_multinomial_UNCLUSTERED.docx",
  clustered  = FALSE,
  verify_ses = FALSE
)

## Carry frequency, gun owners (clustered)
run_multinom_models_to_word(
  dv         = "carry_freq_cat",
  data       = dat_gunowners_processed,
  file_name  = "output/models_carry_freq_multinomial_CLUSTERED.docx",
  clustered  = TRUE,
  verify_ses = FALSE
)

## Carry frequency, gun owners, no small-N states (clustered)
run_multinom_models_to_word(
  dv         = "carry_freq_cat",
  data       = dat_gunowners_processed_nomin,
  file_name  = "output/models_carry_freq_multinomial_CLUSTERED_nomin.docx",
  clustered  = TRUE,
  verify_ses = FALSE
)

## Carry type, gun owners (clustered)
run_multinom_models_to_word(
  dv         = "carry_type_cat",
  data       = dat_gunowners_processed,
  file_name  = "output/models_carry_type_multinomial_CLUSTERED.docx",
  clustered  = TRUE,
  verify_ses = FALSE
)

## Carry type, gun owners, no small-N states (clustered)
run_multinom_models_to_word(
  dv         = "carry_type_cat",
  data       = dat_gunowners_processed_nomin,
  file_name  = "output/models_carry_type_multinomial_CLUSTERED_nomin.docx",
  clustered  = TRUE,
  verify_ses = FALSE
)

#### Play Lindsey Buckingham saying "organic" when everything is done ####
play(load.wave("Organic.wav"))
```

```{r Model Doc Creation | Purchase Plan | All | Controlling for Gun Ownership}
# Separate chunk because we're changing the controls
# Use gun ownership variants from shared specs
controls_formula <- paste(control_vars_gunown, collapse = " + ")
var_labels <- var_labels_gunown
var_order <- var_order_gunown

# WEIGHTED purchase_plan
run_logit_models_to_word(
  dv         = "purchase_plan",
  data       = dat_all_processed,
  weight_var = "wgtpi",
  file_name  = "output/clustered_models_purchase_plan_WEIGHTED.docx"
)

# UNWEIGHTED purchase_plan
run_logit_models_to_word(
  dv         = "purchase_plan",
  data       = dat_all_processed,
  weight_var = NULL,
  file_name  = "output/clustered_models_purchase_plan_UNWEIGHTED.docx"
)

# Play Lindsey Buckingham saying "organic" when the document is done
play(load.wave("Organic.wav"))

```

Note on estimation approach for weighted models:
We use survey::svyglm() to correctly handle probability weights (wgtpi) 
with state-level clustering. This approach properly accounts for both the 
sampling weights and within-state correlation. Comparison with glm() + 
vcovCL() shows materially equivalent results, but the survey package is 
preferred as it correctly treats weights as inverse selection probabilities 
rather than frequency weights

***Random intercept models***
Appendix I
```{r Document Creation Setup}
#### SETUP (binary-only workflow) ####
# Note: control_vars, state_sets, model_titles, var_labels, and var_order
# are defined in the "Shared Analysis Specifications" chunk

# Binary DVs we plan to use
allowed_binary_dvs <- c("carry", "open_carry", "gun_home","gun_own","purchase_plan")

# Build controls formula for this chunk
individual_controls <- paste(control_vars, collapse = " + ")

# Cell formatter
fmt_coef_se <- function(est, se) {
  ifelse(is.na(est), "", paste0(sprintf("%.2f", est), "\n(", sprintf("%.2f", se), ")"))
}

#### MODEL & EXPORT HELPERS ####

# Build model formulas for one DV (all with random intercept for state)
build_formulas_for_dv <- function(dv) {
  f0 <- as.formula(glue("{dv} ~ cc_no_permit + (1 | state)"))
  f1 <- as.formula(glue("{dv} ~ cc_no_permit + {individual_controls} + (1 | state)"))
  f_state <- lapply(state_sets, function(ss)
    as.formula(glue("{dv} ~ cc_no_permit + {individual_controls} + {ss} + (1 | state)"))
  )
  c(list(f0, f1), f_state)
}

# Always fit GLMM (binary DV): logit with random intercept for state
fit_me <- function(fml, data) {
  glmer(fml, data = data, family = binomial(),
        control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
}

# Tidy GLMM fixed effects; report OR and delta-method SE(OR) ≈ OR * SE(log-odds)
tidy_glmm_or <- function(mod) {
  tt <- broom.mixed::tidy(mod, effects = "fixed", conf.int = FALSE)
  tt <- tt %>%
    mutate(
      OR     = exp(estimate),
      se_or  = OR * std.error,
      p      = if ("p.value" %in% names(.)) p.value else
                 tryCatch(2 * pnorm(-abs(statistic)), error = function(e) NA_real_)
    )
  tt
}

# Assemble the wide table for one DV and return both the table & models
assemble_table_for_dv <- function(dv, data) {
  stopifnot(dv %in% allowed_binary_dvs)
  fmls <- build_formulas_for_dv(dv)

  mods <- map(fmls, fit_me, data = data)
  names(mods) <- model_titles[seq_along(mods)]

  # Tidy each model and reshape into [var, <title>_coef_se, <title>_stars]
  pieces <- imap(mods, function(m, nm) {
    ti <- tidy_glmm_or(m) %>%
      filter(term != "(Intercept)") %>%
      mutate(
        var     = var_labels[term] %||% term,
        coef_se = fmt_coef_se(OR, se_or),
        stars   = sig_stars(p)
      ) %>%
      dplyr::select(var, coef_se, stars)
    names(ti)[2:3] <- c(paste0(nm, "_coef_se"), paste0(nm, "_stars"))
    ti
  })

  wide <- reduce(pieces, full_join, by = "var")
  wide[is.na(wide)] <- ""

  list(table = wide, models = mods)
}

# Compact flextable builder (expects columns: Variable + repeated pairs)
build_flextable_for_dv <- function(dv, tab2) {

  stopifnot("Variable" %in% names(tab2))

  coef_cols  <- grep("_coef_se$",  names(tab2))
  stars_cols <- grep("_stars$",    names(tab2))
  stopifnot(length(coef_cols) == length(stars_cols))

  cols_no_var  <- names(tab2)[-1]
  lower_labels <- ifelse(
    endsWith(cols_no_var, "_coef_se"),
    "Coef (SE)",
    "Sig."
  )

  groups  <- sub("_(coef_se|stars)$", "", cols_no_var)
  grp_rle <- rle(groups)

  ft <- flextable::flextable(tab2)
  ft <- flextable::theme_booktabs(ft)

  # bottom header row labels
  for (i in seq_along(cols_no_var)) {
    ft <- flextable::set_header_labels(
      ft,
      .labels = stats::setNames(lower_labels[i], cols_no_var[i])
    )
  }
  
  # top header row (model titles)
  ft <- flextable::add_header_row(
    ft,
    values    = c(stringr::str_to_title(dv), grp_rle$values),
    colwidths = c(1, grp_rle$lengths)
  )
  ft <- flextable::merge_h(ft, part = "header")
  
  # remove redundant second header row
  ft <- flextable::delete_rows(ft, i = 2, part = "header")
  
  # blank star-column headers
  star_cols <- grep("_stars$", names(tab2))
  ft <- flextable::set_header_labels(
    ft,
    values = setNames(rep("", length(star_cols)), names(tab2)[star_cols])
  )
  
  # replace top-left header cell
  ft <- flextable::set_header_labels(
    ft,
    values = c(Variable = "Term")
  )

  # ---- Alignment ----
  ft <- flextable::align(ft, j = 1,  align = "left",   part = "all")
  ft <- flextable::align(ft, j = -1, align = "center", part = "all")
  ft <- flextable::valign(ft,        val   = "center", part = "all")

  # ---- FIXED LAYOUT (prevents Word reflow) ----
  ft <- flextable::set_table_properties(
    ft,
    layout = "fixed",
    width  = 1
  )

  # ---- Font defaults ----
  ft <- flextable::font(ft, fontname = "Times New Roman", part = "all")
  ft <- flextable::fontsize(ft, size = 10, part = "all")

  # ---- Padding ----
  ft <- flextable::padding(ft, padding = 2, part = "all")

  # ---- Column widths (MATCH clustered logit + multinomial) ----
  ft <- flextable::width(ft, j = 1,          width = 1.40)  # Variable
  ft <- flextable::width(ft, j = coef_cols,  width = 0.85)  # Coef(SE)
  ft <- flextable::width(ft, j = stars_cols, width = 0.35)  # Sig.

  # ---- Rules ----
  thin <- officer::fp_border(color = "#999999", width = 0.75)
  ft <- flextable::border_remove(ft)
  ft <- flextable::hline_top(ft,    border = thin, part = "header")
  ft <- flextable::hline_bottom(ft, border = thin, part = "header")
  ft <- flextable::hline_bottom(ft, border = thin, part = "body")
  ft <- flextable::bold(ft, part = "header")

  ft
}

```

```{r Model Doc Creation | Gun Ownership | Unweighted}
# Choose DV (from allowed_binary_dvs list)
dv <- "gun_own"

# filter data
matched_data <- make_matched_sample(
  dv = dv,
  data = dat_all_processed,
  individual_controls = individual_controls,
  state_sets = state_sets
)

message("Fitting mixed-effects models for DV = ", dv)
assembled <- assemble_table_for_dv(dv, matched_data)
tab <- assembled$table
mods <- assembled$models

# Rename columns to "Variable", then "<label>_coef_se", "<label>_stars"
#    using the model names we actually have (Base, Ind., set1..)
stopifnot(ncol(tab) > 0)
k <- length(mods)
model_labels <- names(mods)
new_names <- c("Variable", as.vector(rbind(
  paste0(model_labels, "_coef_se"),
  paste0(model_labels, "_stars")
)))
names(tab) <- new_names

# Footer with state variance, N, and AIC (for each model)
gl_list <- lapply(mods, function(m) {
  gl <- tryCatch(broom.mixed::glance(m), error = function(e) NULL)
  vc <- tryCatch(as.data.frame(VarCorr(m)), error = function(e) NULL)
  state_var <- if (!is.null(vc)) vc$vcov[vc$grp == "state"] else NA
  c(
    StateVar = state_var,
    N   = if (!is.null(gl) && !is.null(gl$nobs)) gl$nobs else nobs(m),
    AIC = if (!is.null(gl) && !is.null(gl$AIC))  gl$AIC  else AIC(m)
  )
})
gl_df <- do.call(rbind, gl_list)

footer_rows <- list()
for (stat_name in c("State variance", "AIC")) {
  row <- setNames(as.list(rep("", length(new_names))), new_names)
  row$Variable <- stat_name
  for (j in seq_len(k)) {
    val <- gl_df[j, switch(stat_name,
                           "State variance" = "StateVar",
                           "AIC"            = "AIC")]
    row[[ paste0(model_labels[j], "_coef_se") ]] <-
      if (stat_name == "AIC") formatC(as.numeric(val), format = "f", digits = 1)
      else if (stat_name == "State variance") formatC(as.numeric(val), format = "f", digits = 3)
    row[[ paste0(model_labels[j], "_stars") ]] <- ""
  }
  footer_rows[[length(footer_rows) + 1]] <- as.data.frame(row, check.names = FALSE)
}
footer <- bind_rows(footer_rows)

tab2 <- bind_rows(tab, footer)
tab2[is.na(tab2)] <- ""

# Build flextable & export to Word
ft <- build_flextable_for_dv(dv, tab2)

doc_path <- paste0("output/models_permit_", dv, "_mixed.docx")
doc <- officer::read_docx()
doc <- officer::body_add_par(doc, stringr::str_to_title(dv), style = "heading 1")
doc <- flextable::body_add_flextable(doc, ft)
print(doc, target = doc_path)
message("Wrote: ", doc_path)

# Play Lindsey Buckingham saying "organic" when the document is done
play(load.wave("Organic.wav"))
```

```{r Model Doc Creation | Gun in Home | Unweighted}
# Choose DV (from allowed_binary_dvs list)
dv <- "gun_home"

# filter data
matched_data <- make_matched_sample(
  dv = dv,
  data = dat_all_processed,
  individual_controls = individual_controls,
  state_sets = state_sets
)

message("Fitting mixed-effects models for DV = ", dv)
assembled <- assemble_table_for_dv(dv, matched_data)
tab <- assembled$table
mods <- assembled$models

# Rename columns to "Variable", then "<label>_coef_se", "<label>_stars"
# using the model names we actually have (Base, Ind., set1..)
stopifnot(ncol(tab) > 0)
k <- length(mods)
model_labels <- names(mods)
new_names <- c("Variable", as.vector(rbind(
  paste0(model_labels, "_coef_se"),
  paste0(model_labels, "_stars")
)))
names(tab) <- new_names

# Footer with state variance, N, and AIC (for each model)
gl_list <- lapply(mods, function(m) {
  gl <- tryCatch(broom.mixed::glance(m), error = function(e) NULL)
  vc <- tryCatch(as.data.frame(VarCorr(m)), error = function(e) NULL)
  state_var <- if (!is.null(vc)) vc$vcov[vc$grp == "state"] else NA
  c(
    StateVar = state_var,
    N   = if (!is.null(gl) && !is.null(gl$nobs)) gl$nobs else nobs(m),
    AIC = if (!is.null(gl) && !is.null(gl$AIC))  gl$AIC  else AIC(m)
  )
})
gl_df <- do.call(rbind, gl_list)

footer_rows <- list()
for (stat_name in c("State variance", "AIC")) {
  row <- setNames(as.list(rep("", length(new_names))), new_names)
  row$Variable <- stat_name
  for (j in seq_len(k)) {
    val <- gl_df[j, switch(stat_name,
                           "State variance" = "StateVar",
                           "AIC"            = "AIC")]
    row[[ paste0(model_labels[j], "_coef_se") ]] <-
      if (stat_name == "AIC") formatC(as.numeric(val), format = "f", digits = 1)
      else if (stat_name == "State variance") formatC(as.numeric(val), format = "f", digits = 3)
    row[[ paste0(model_labels[j], "_stars") ]] <- ""
  }
  footer_rows[[length(footer_rows) + 1]] <- as.data.frame(row, check.names = FALSE)
}
footer <- bind_rows(footer_rows)

tab2 <- bind_rows(tab, footer)
tab2[is.na(tab2)] <- ""

# Build flextable & export to Word
ft <- build_flextable_for_dv(dv, tab2)

doc_path <- paste0("output/models_permit_", dv, "_mixed.docx")
doc <- officer::read_docx()
doc <- officer::body_add_par(doc, stringr::str_to_title(dv), style = "heading 1")
doc <- flextable::body_add_flextable(doc, ft)
print(doc, target = doc_path)
message("Wrote: ", doc_path)

# Play Lindsey Buckingham saying "organic" when the document is done
play(load.wave("Organic.wav"))
```

```{r Model Doc Creation | Carry | Unweighted | Gun Owners}
# Choose DV (from allowed_binary_dvs list)
dv <- "carry"

# filter data
matched_data <- make_matched_sample(
  dv = dv,
  data = dat_gunowners_processed,
  individual_controls = individual_controls,
  state_sets = state_sets
)

message("Fitting mixed-effects models for DV = ", dv)
assembled <- assemble_table_for_dv(dv, matched_data)
tab <- assembled$table
mods <- assembled$models

# Rename columns to "Variable", then "<label>_coef_se", "<label>_stars"
# using the model names we actually have (Base, Ind., set1..)
stopifnot(ncol(tab) > 0)
k <- length(mods)
model_labels <- names(mods)
new_names <- c("Variable", as.vector(rbind(
  paste0(model_labels, "_coef_se"),
  paste0(model_labels, "_stars")
)))
names(tab) <- new_names

# Footer with state variance, N, and AIC (for each model)
gl_list <- lapply(mods, function(m) {
  gl <- tryCatch(broom.mixed::glance(m), error = function(e) NULL)
  vc <- tryCatch(as.data.frame(VarCorr(m)), error = function(e) NULL)
  state_var <- if (!is.null(vc)) vc$vcov[vc$grp == "state"] else NA
  c(
    StateVar = state_var,
    N   = if (!is.null(gl) && !is.null(gl$nobs)) gl$nobs else nobs(m),
    AIC = if (!is.null(gl) && !is.null(gl$AIC))  gl$AIC  else AIC(m)
  )
})
gl_df <- do.call(rbind, gl_list)

footer_rows <- list()
for (stat_name in c("State variance", "AIC")) {
  row <- setNames(as.list(rep("", length(new_names))), new_names)
  row$Variable <- stat_name
  for (j in seq_len(k)) {
    val <- gl_df[j, switch(stat_name,
                           "State variance" = "StateVar",
                           "AIC"            = "AIC")]
    row[[ paste0(model_labels[j], "_coef_se") ]] <-
      if (stat_name == "AIC") formatC(as.numeric(val), format = "f", digits = 1)
      else if (stat_name == "State variance") formatC(as.numeric(val), format = "f", digits = 3)
    row[[ paste0(model_labels[j], "_stars") ]] <- ""
  }
  footer_rows[[length(footer_rows) + 1]] <- as.data.frame(row, check.names = FALSE)
}
footer <- bind_rows(footer_rows)

tab2 <- bind_rows(tab, footer)
tab2[is.na(tab2)] <- ""

# Build flextable & export to Word
ft <- build_flextable_for_dv(dv, tab2)

doc_path <- paste0("output/models_permit_", dv, "_mixed.docx")
doc <- officer::read_docx()
doc <- officer::body_add_par(doc, stringr::str_to_title(dv), style = "heading 1")
doc <- flextable::body_add_flextable(doc, ft)
print(doc, target = doc_path)
message("Wrote: ", doc_path)

# Play Lindsey Buckingham saying "organic" when the document is done
play(load.wave("Organic.wav"))
```

```{r Model Doc Creation | Open Carry | Unweighted | Gun Owners}
# Choose DV (from allowed_binary_dvs list)
dv <- "open_carry"

# filter data
matched_data <- make_matched_sample(
  dv = dv,
  data = dat_gunowners_processed,
  individual_controls = individual_controls,
  state_sets = state_sets
)

message("Fitting mixed-effects models for DV = ", dv)
assembled <- assemble_table_for_dv(dv, matched_data)
tab <- assembled$table
mods <- assembled$models

# Rename columns to "Variable", then "<label>_coef_se", "<label>_stars"
# using the model names we actually have (Base, Ind., set1..)
stopifnot(ncol(tab) > 0)
k <- length(mods)
model_labels <- names(mods)
new_names <- c("Variable", as.vector(rbind(
  paste0(model_labels, "_coef_se"),
  paste0(model_labels, "_stars")
)))
names(tab) <- new_names

# Footer with state variance, N, and AIC (for each model)
gl_list <- lapply(mods, function(m) {
  gl <- tryCatch(broom.mixed::glance(m), error = function(e) NULL)
  vc <- tryCatch(as.data.frame(VarCorr(m)), error = function(e) NULL)
  state_var <- if (!is.null(vc)) vc$vcov[vc$grp == "state"] else NA
  c(
    StateVar = state_var,
    N   = if (!is.null(gl) && !is.null(gl$nobs)) gl$nobs else nobs(m),
    AIC = if (!is.null(gl) && !is.null(gl$AIC))  gl$AIC  else AIC(m)
  )
})
gl_df <- do.call(rbind, gl_list)

footer_rows <- list()
for (stat_name in c("State variance", "AIC")) {
  row <- setNames(as.list(rep("", length(new_names))), new_names)
  row$Variable <- stat_name
  for (j in seq_len(k)) {
    val <- gl_df[j, switch(stat_name,
                           "State variance" = "StateVar",
                           "AIC"            = "AIC")]
    row[[ paste0(model_labels[j], "_coef_se") ]] <-
      if (stat_name == "AIC") formatC(as.numeric(val), format = "f", digits = 1)
      else if (stat_name == "State variance") formatC(as.numeric(val), format = "f", digits = 3)
    row[[ paste0(model_labels[j], "_stars") ]] <- ""
  }
  footer_rows[[length(footer_rows) + 1]] <- as.data.frame(row, check.names = FALSE)
}
footer <- bind_rows(footer_rows)

tab2 <- bind_rows(tab, footer)
tab2[is.na(tab2)] <- ""

# Build flextable & export to Word
ft <- build_flextable_for_dv(dv, tab2)

doc_path <- paste0("output/models_permit_", dv, "_mixed.docx")
doc <- officer::read_docx()
doc <- officer::body_add_par(doc, stringr::str_to_title(dv), style = "heading 1")
doc <- flextable::body_add_flextable(doc, ft)
print(doc, target = doc_path)
message("Wrote: ", doc_path)

# Play Lindsey Buckingham saying "organic" when the document is done
play(load.wave("Organic.wav"))
```

```{r Model Doc Creation | Purchase Plan | Unweighted | Controlling for gun ownership}

#### Use gun ownership variants from shared specs ####
individual_controls <- paste(control_vars_gunown, collapse = " + ")
var_labels <- var_labels_gunown

# Choose DV (from allowed_binary_dvs list)
dv <- "purchase_plan"

# filter data
matched_data <- make_matched_sample(
  dv = dv,
  data = dat_all_processed,
  individual_controls = individual_controls,
  state_sets = state_sets
)

message("Fitting mixed-effects models for DV = ", dv)
assembled <- assemble_table_for_dv(dv, matched_data)
tab <- assembled$table
mods <- assembled$models

# 2) Rename columns to "Variable", then "<label>_coef_se", "<label>_stars"
#    using the model names we actually have (Base, Ind., set1..)
stopifnot(ncol(tab) > 0)
k <- length(mods)
model_labels <- names(mods)
new_names <- c("Variable", as.vector(rbind(
  paste0(model_labels, "_coef_se"),
  paste0(model_labels, "_stars")
)))
names(tab) <- new_names

# 3) Footer with state variance, N, and AIC (for each model)
gl_list <- lapply(mods, function(m) {
  gl <- tryCatch(broom.mixed::glance(m), error = function(e) NULL)
  vc <- tryCatch(as.data.frame(VarCorr(m)), error = function(e) NULL)
  state_var <- if (!is.null(vc)) vc$vcov[vc$grp == "state"] else NA
  c(
    StateVar = state_var,
    N   = if (!is.null(gl) && !is.null(gl$nobs)) gl$nobs else nobs(m),
    AIC = if (!is.null(gl) && !is.null(gl$AIC))  gl$AIC  else AIC(m)
  )
})
gl_df <- do.call(rbind, gl_list)

footer_rows <- list()
for (stat_name in c("State variance", "AIC")) {
  row <- setNames(as.list(rep("", length(new_names))), new_names)
  row$Variable <- stat_name
  for (j in seq_len(k)) {
    val <- gl_df[j, switch(stat_name,
                           "State variance" = "StateVar",
                           "AIC"            = "AIC")]
    row[[ paste0(model_labels[j], "_coef_se") ]] <-
      if (stat_name == "AIC") formatC(as.numeric(val), format = "f", digits = 1)
      else if (stat_name == "State variance") formatC(as.numeric(val), format = "f", digits = 3)
    row[[ paste0(model_labels[j], "_stars") ]] <- ""
  }
  footer_rows[[length(footer_rows) + 1]] <- as.data.frame(row, check.names = FALSE)
}
footer <- bind_rows(footer_rows)

tab2 <- bind_rows(tab, footer)
tab2[is.na(tab2)] <- ""

# 4) Build flextable & export to Word
ft <- build_flextable_for_dv(dv, tab2)

doc_path <- paste0("output/models_permit_", dv, "_mixed.docx")
doc <- officer::read_docx()
doc <- officer::body_add_par(doc, stringr::str_to_title(dv), style = "heading 1")
doc <- flextable::body_add_flextable(doc, ft)
print(doc, target = doc_path)
message("Wrote: ", doc_path)

# Play Lindsey Buckingham saying "organic" when the document is done
play(load.wave("Organic.wav"))
```

```{r See a given model}
assembled <- assemble_table_for_dv("gun_own", matched_data)
mods <- assembled$models
names(mods)
# [1] "Base (RI only)" "Ind. controls" "set1" "set2" "set3" "set4" "set5"

# RENAME MODELS FROM NAMES
m1 <- mods[["set4"]]
m2  <- mods[["Ind. controls"]]

summary(m1)        # fixed effects, z, p; random-intercept SD
lme4::fixef(m1)    # fixed effects only
lme4::ranef(m1)    # random effects (BLUPs) by state
lme4::VarCorr(m1)  # random-effects variance/covariance
AIC(m1); nobs(m1)  # fit stats

# ORs for all fixed effects
exp(lme4::fixef(m1))

# 95% CI for ORs (Wald)
coefs <- lme4::fixef(m1)
ses   <- sqrt(diag(vcov(m1)))
OR    <- exp(coefs)
lo    <- exp(coefs - 1.96 * ses)
hi    <- exp(coefs + 1.96 * ses)
cbind(OR, lo, hi)

# fixed effects with OR and an approximate SE(OR)
broom.mixed::tidy(m1, effects = "fixed") |>
  dplyr::mutate(OR = exp(estimate),
                se_or = OR * std.error)

# one-row model summary (AIC, nobs, etc.)
broom.mixed::glance(m1)

# compare model fit
AIC(m1, m2)               # smaller = better (penalized for complexity)
anova(m1, m2, test = "Chisq")  # LR test (nested models)
```
