---
output: console
---

Weight Creation Script
ACS 2024
Variables:
- Numeric: gender_num, age_num, educ_num, inc5_num, pid3_num, race_eth_reg_num
- Factor: gender_cat, age_cat, educ_cat, inc5_cat, pid3_cat, race_eth_reg_cat

```{r Notes}

# Worth noting that the weighting here we are doing is on an op-in panel. This isn't a deal where we're weighting to correct for sampling probabilities. We have no sampling frame because we have no idea why any given person ended up in the sample. We're essentially weighting as an act of faith to make it closer to something representative of the true population

# Reasoning behind the raking variables:
  # Gender, age, education are typical raking vars.
  # Income - opt-in panels tend to skew more middle-income, and potentially relevant for outcomes
  # Race x Region is more precise to the racial imbalances across the country and is theorized to be important especially in terms of gun culture
  # PID is relevant to the outcomes of interest—only problematic if the lurking variable that leads to under/over-representation of certain political groups also affects our outcomes of interest

# Can't correct for the fact that people in opt-in panels likely differ from the general population in important ways

# Notes with a bit more info that I didn't want to write down from Claude:
# RAKING OVERVIEW

# Raking (iterative proportional fitting) adjusts each respondent's weight so
# that the weighted sample matches known population targets on each variable
# margin. It works by cycling through variables one at a time, rescaling
# weights to hit each target, and repeating until all margins converge.
#
# We rake on six variables:
#   - Gender, age, education, income, race x region: standard demographic
#     corrections for the known biases of opt-in online panels like Lucid,
#     which tend to skew higher-education, middle-income, and regionally
#     unrepresentative
#   - Party ID: included because partisan composition strongly predicts
#     attitudes toward political violence -- the primary outcome of interest --
#     so an imbalanced sample on PID would directly bias key estimates.
#     The tradeoff is that PID is not a stable demographic characteristic;
#     if the selection process that causes certain partisan groups to be
#     over/underrepresented in Lucid is itself related to political violence
#     attitudes beyond what PID captures, raking on PID corrects the surface
#     imbalance but not the underlying selection bias. On balance, inclusion
#     is preferable given how strongly partisan our outcomes are.
#   - Race x region interaction rather than race and region separately:
#     racial composition varies dramatically by region (e.g., the South is
#     heavily Black, the West heavily Hispanic). Raking on independent margins
#     can produce a sample that hits both marginals but still misrepresents
#     Black Southerners or Hispanic Westerners. The interaction forces the
#     joint distribution to match the population.
#
# WHEN RAKING ON A VARIABLE IS PROBLEMATIC:
#   Raking assumes the variable is a stable population characteristic with
#   known targets and that any over/underrepresentation is incidental to the
#   sampling process rather than related to your outcomes. PID is the main
#   case where this assumption is debatable (see above). You would NOT want
#   to rake on an outcome variable or a variable causally downstream of your
#   outcomes -- that would introduce bias rather than reduce it.
```

```{r Load packages}
# Library setup
library(data.table) # For fread() -- query a CSV directly without loading it into memory
library(survey) # For survey::rake
library(tidyverse) # I rely on the %>% operator from {tidyverse}
```

```{r Load acs data, eval=FALSE}
# # Originally used ACS PUMS to derive population targets, but the files are huge and it takes a lot of memory so can skip to the chunks below
# # We get the pid and income estimates from other sources anyway
# 
# # Census data is huge, so selectively read in the variables we need
# a <- fread(
#   "~/Library/CloudStorage/GoogleDrive-pauleteas@gmail.com/My Drive/Research/PROJECTS/Bedrock/weight_creation/csv_pus/psam_pusa.csv",
#   select = c("HISP", "AGEP", "REGION", "RAC1P", "SCHL", "SEX", "PWGTP")
# )
# b <- fread(
#   "~/Library/CloudStorage/GoogleDrive-pauleteas@gmail.com/My Drive/Research/PROJECTS/Bedrock/weight_creation/csv_pus/psam_pusb.csv",
#   select = c("HISP", "AGEP", "REGION", "RAC1P", "SCHL", "SEX", "PWGTP")
# )
# 
# # Combine the csvs
# acs <- rbind(a, b)
# rm(a, b) # Clear the individual objects to save on memory
# acs <- subset(acs, acs$AGEP > 17) # restrict data to adults
# head(acs)
# 
# summary(acs$PWGTP) # This is the weights that ACS uses
# # normal to have giant range for these because it's weighted to represent literally the whole population, not a sample of the population. So like on average, each participant represents 96 American adults
# 
# # Recode ACS variables to match survey coding
# acs$hispanic[acs$HISP != 1] <- 1  # Hispanic
# acs$hispanic[acs$HISP == 1] <- 2  # Non-Hispanic
# 
# acs$sex[acs$SEX == 1] <- 1 # Male
# acs$sex[acs$SEX == 2] <- 2 # Female
# 
# acs$region[acs$REGION == 1] <- 1 # Northeast
# acs$region[acs$REGION == 2] <- 2 # Midwest
# acs$region[acs$REGION == 3] <- 3 # South
# acs$region[acs$REGION == 4] <- 4 # West
# 
# acs$age4[acs$AGEP < 30] <- 1  # 18-29
# acs$age4[acs$AGEP >= 30 & acs$AGEP < 45] <- 2  # 30-44
# acs$age4[acs$AGEP >= 45 & acs$AGEP < 65] <- 3  # 45-64
# acs$age4[acs$AGEP >= 65] <- 4  # 65+
# 
# acs$educ4[acs$SCHL <= 17] <- 1 # HS or less (includes GED)
# acs$educ4[acs$SCHL >= 18 & acs$SCHL <= 20] <- 2 # Some college + Associate's
# acs$educ4[acs$SCHL == 21] <- 3 # Bachelor's degree
# acs$educ4[acs$SCHL >= 22] <- 4 # Graduate degree
# 
# acs$race4 <- 4  # Other (default)
# acs$race4[acs$RAC1P == 1 & acs$hispanic == 2] <- 1  # White non-Hispanic
# acs$race4[acs$RAC1P == 2 & acs$hispanic == 2] <- 2  # Black non-Hispanic
# acs$race4[acs$hispanic == 1] <- 3  # Hispanic (any race)
# 
# # Race x Region interaction
# acs$race_eth_reg <- acs$race4 + (acs$region - 1) * 4
# 
# # Create labeled factor versions for clarity
# # _num = numeric (sex, age4, etc.), _cat = factors with labels
# acs <- acs %>%
#   dplyr::mutate(
#     sex_cat = factor(sex, levels = 1:2, labels = c("Male", "Female")),
#     age4_cat = factor(age4, levels = 1:4, labels = c("18-29", "30-44", "45-64", "65+")),
#     educ4_cat = factor(educ4, levels = 1:4, labels = c("HS or less", "Some college", "College", "Graduate")),
#     race4_cat = factor(race4, levels = 1:4, labels = c("White", "Black", "Hispanic", "Other")),
#     region_cat = factor(region, levels = 1:4, labels = c("Northeast", "Midwest", "South", "West")),
#     race_eth_reg_cat = factor(race_eth_reg, levels = 1:16,
#             labels = c("NE White", "NE Black", "NE Hispanic", "NE Other",
#                        "MW White", "MW Black", "MW Hispanic", "MW Other",
#                        "S White", "S Black", "S Hispanic", "S Other",
#                        "W White", "W Black", "W Hispanic", "W Other"))
#   )
```

```{r Calculate acs targets, eval=FALSE}
# # Calculate weighted population proportions from ACS using PWGTP
# # These become our "target" proportions for raking
# # Using labeled factors for easier interpretation
# 
# # Gender targets
# acs %>%
#   filter(!is.na(sex_cat)) %>%
#   group_by(sex_cat) %>%
#   summarise(weighted_n = sum(PWGTP, na.rm = TRUE)) %>%
#   mutate(prop = weighted_n / sum(weighted_n) * 100)
# 
# # Age targets (4 categories)
# acs %>%
#   filter(!is.na(age4_cat)) %>%
#   group_by(age4_cat) %>%
#   summarise(weighted_n = sum(PWGTP, na.rm = TRUE)) %>%
#   mutate(prop = weighted_n / sum(weighted_n) * 100)
# 
# # Race targets (4 categories)
# acs %>%
#   filter(!is.na(race4_cat)) %>%
#   group_by(race4_cat) %>%
#   summarise(weighted_n = sum(PWGTP, na.rm = TRUE)) %>%
#   mutate(prop = weighted_n / sum(weighted_n) * 100)
# 
# # Education targets (4 categories)
# acs %>%
#   filter(!is.na(educ4_cat)) %>%
#   group_by(educ4_cat) %>%
#   summarise(weighted_n = sum(PWGTP, na.rm = TRUE)) %>%
#   mutate(prop = weighted_n / sum(weighted_n) * 100)
# 
# # Region targets
# acs %>%
#   filter(!is.na(region_cat)) %>%
#   group_by(region_cat) %>%
#   summarise(weighted_n = sum(PWGTP, na.rm = TRUE)) %>%
#   mutate(prop = weighted_n / sum(weighted_n) * 100)
# 
# # Race x Region targets (16 categories)
# acs %>%
#   filter(!is.na(race_eth_reg_cat)) %>%
#   group_by(race_eth_reg_cat) %>%
#   summarise(weighted_n = sum(PWGTP, na.rm = TRUE)) %>%
#   mutate(prop = weighted_n / sum(weighted_n) * 100)
# 
# rm(acs) # Clear ACS to save memory
```

```{r Calculate income targets from Census table}
# NOTE: Income is NOT available in the ACS PUMS raw data files we used above.
# Instead, we calculate income distribution targets from the Census Bureau's
# Current Population Survey (CPS) HINC-06 table: "Income Distribution to $250,000 or More for Households"
# Source: https://www.census.gov/data/tables/2025/demo/cps/hinc-06.html
# Data file: income.csv (extracted from HINC-06 table for 2024)

# Read the Census income distribution table
income <- read.csv("income.csv", header = FALSE, stringsAsFactors = FALSE)
names(income) <- c("bracket", "households")

# Clean the data - remove commas and convert to numeric
income$households <- as.numeric(gsub(",", "", income$households))

# Get total households (first row)
total <- income$households[1]
total
head(income)

# Calculate counts for each of our 5 income categories
# Up to $29,999 (rows 2-7): Under $5k through $25-30k
under_30k <- sum(income$households[2:7])

# $30,000-$59,999 (rows 8-13): $30-35k through $55-60k
inc_30_60k <- sum(income$households[8:13])

# $60,000-$99,999 (rows 14-21): $60-65k through $95-100k
inc_60_100k <- sum(income$households[14:21])

# $100,000-$149,999 (rows 22-31): $100-105k through $145-150k
inc_100_150k <- sum(income$households[22:31])

# $150,000+ (rows 32-43): $150-155k through $250k+
inc_150k_plus <- sum(income$households[32:43])

# Calculate percentages
round((under_30k / total) * 100, 1)
round((inc_30_60k / total) * 100, 1)
round((inc_60_100k / total) * 100, 1)
round((inc_100_150k / total) * 100, 1)
round((inc_150k_plus / total) * 100, 1)
```

```{r Define targets}
# Using text labels that match our factor levels for clarity
target_gender <- c(
  "Male" = 49.0,
  "Female" = 51.0
)

target_age <- c(
  "18-29" = 20.0,
  "30-44" = 26.2,
  "45-64" = 30.9,
  "65+" = 22.9
)

target_educ <- c(
  "HS or less" = 37.3,
  "Some college" = 28.5,
  "College" = 21.0,
  "Graduate" = 13.2
)

# income targets from Census CPS HINC-06 table
target_inc5 <- c(
  "Up to $29,999" = 16.8,
  "$30,000-$59,999" = 19.6,
  "$60,000-$99,999" = 20.8,
  "$100,000-$149,999" = 16.7,
  "$150,000+" = 26.1
)

# targets from Gallup 2024: https://news.gallup.com/poll/15370/party-affiliation.aspx
target_pid3 <- c(
  "Democrat" = 45,
  "Independent" = 9,
  "Republican" = 46
)

# Race x Region interaction (16 categories)
target_race_eth_reg <- c(
  "NE White" = 10.9,
  "NE Black" = 1.78,
  "NE Hispanic" = 2.56,
  "NE Other" = 2.04,
  "MW White" = 15.1,
  "MW Black" = 1.92,
  "MW Hispanic" = 1.69,
  "MW Other" = 1.65,
  "S White" = 21.4,
  "S Black" = 6.82,
  "S Hispanic" = 7.16,
  "S Other" = 3.33,
  "W White" = 11.5,
  "W Black" = 1.02,
  "W Hispanic" = 6.71,
  "W Other" = 4.3
)

# Targets should sum to ~100
sum(target_gender)
sum(target_age)
sum(target_educ)
sum(target_inc5)
sum(target_pid3)
sum(target_race_eth_reg)
```

```{r Load survey data}
dat <- read.csv(
  "data/data_unweighted.csv",
  header = TRUE,
  stringsAsFactors = FALSE
)

# Remove non-Lucid data
dat <- dat %>% filter(gc == 1)

# checks
nrow(dat)
table(dat$region) # From Lucid/Cint
table(dat$Q2) # Age
table(dat$Q3) # Gender
table(dat$Q4) # Hispanic
table(dat$Q5) # Race
table(dat$standard_education_us) # From Lucid/Cint
table(dat$standard_hhi) # From Lucid/Cint
table(dat$Q10) # Party ID
table(dat$Q13) # Leaner question
```

```{r Recode survey data}
# All starting variables are numeric/integers
# Create _num variables for numeric values

# Case ID
dat <- dat %>%
  dplyr::mutate(caseid = dplyr::row_number())

# Gender (1 = Male, 2 = Female)
# Q3 is 1=Female, 2=Male in survey, but need to flip to match ACS (1=Male, 2=Female)
dat <- dat %>%
  dplyr::mutate(
    gender_num = dplyr::case_when(
      Q3 == 1 ~ 2,  # Survey Female (1) -> ACS Female (2)
      Q3 == 2 ~ 1,  # Survey Male (2) -> ACS Male (1)
      TRUE ~ NA_real_
    )
  )
table(dat$gender_num, useNA = "ifany")

# Age (1 = 18-29, 2 = 30-44, 3 = 45-64, 4 = 65+)
# Q2 is 2-5 in survey (1="under 18" was excluded), recode to 1-4
dat <- dat %>%
  dplyr::mutate(
    age_num = dplyr::case_when(
      Q2 == 2 ~ 1,  # 18-29
      Q2 == 3 ~ 2,  # 30-44
      Q2 == 4 ~ 3,  # 45-64
      Q2 == 5 ~ 4,  # 65+
      TRUE ~ NA_real_
    )
  )
table(dat$age_num, useNA = "ifany")

# Race4 (1 = White non-Hispanic, 2 = Black non-Hispanic, 3 = Hispanic, 4 = Other)
# Q4 is Hispanic indicator, Q5 is race
# Need to create race_num based on numeric Q4 and Q5
dat <- dat %>%
  dplyr::mutate(
    Q4_num = as.numeric(Q4),
    Q5_num = as.numeric(Q5),
    race_num = dplyr::case_when(
      Q4_num == 1 ~ 3, # Hispanic (1 = Yes)
      Q4_num == 2 & Q5_num == 1 ~ 1, # White non-Hispanic (2 = No, 1 = White)
      Q4_num == 2 & Q5_num == 2 ~ 2, # Black non-Hispanic (2 = Black)
      Q4_num == 2 ~ 4, # Other non-Hispanic
      TRUE ~ NA_real_
    )
  )
table(dat$race_num, useNA = "ifany")

# Education (1 = HS or less, 2 = Some college, 3 = College, 4 = Grad)
# standard_education_us is already numeric
dat <- dat %>%
  dplyr::mutate(
    edu_num = as.numeric(standard_education_us),
    educ_num = dplyr::case_when(
      edu_num %in% c(1, 2) ~ 1, # HS or less
      edu_num %in% c(3, 4, 5) ~ 2, # Some college + Associate's
      edu_num == 6 ~ 3, # Bachelor's
      edu_num %in% c(7, 8) ~ 4, # Graduate
      TRUE ~ NA_real_
    )
  )
table(dat$educ_num, useNA = "ifany")

# Region (1 = NE, 2 = MW, 3 = S, 4 = W)
# region is already numeric (1, 2, 3, 4)
dat <- dat %>%
  dplyr::mutate(
    region_num = as.numeric(region)
  )
table(dat$region_num, useNA = "ifany")

# Income (5 categories)
# standard_hhi is already numeric, create 5 categories
dat <- dat %>%
  dplyr::mutate(
    standard_hhi_num = as.numeric(standard_hhi),
    inc5_num = dplyr::case_when(
      standard_hhi_num >= 1 & standard_hhi_num <= 4 ~ 1, # Up to $29,999
      standard_hhi_num >= 5 & standard_hhi_num <= 10 ~ 2, # $30,000-$59,999
      standard_hhi_num >= 11 & standard_hhi_num <= 18 ~ 3, # $60,000-$99,999
      standard_hhi_num %in% c(19, 20) ~ 4, # $100,000-$149,999
      standard_hhi_num >= 21 & standard_hhi_num <= 24 ~ 5, # $150,000+
      TRUE ~ NA_real_
    )
  )
table(dat$inc5_num, useNA = "ifany")

# Party ID 3 categories (with leaners)
# 1 = Democrat (including leaners), 2 = Independent, 3 = Republican (including leaners)
# Q10 and Q13 are already numeric
dat <- dat %>%
  dplyr::mutate(
    Q10_num = as.numeric(Q10),
    Q13_num = as.numeric(Q13),
    pid3_num = dplyr::case_when(
      Q10_num == 1 ~ 1, # Democrat
      Q13_num == 1 ~ 1, # Lean Democrat
      Q13_num == 3 ~ 2, # True Independent
      Q10_num == 2 ~ 3, # Republican
      Q13_num == 2 ~ 3, # Lean Republican
      TRUE ~ NA_real_
    )
  )
table(dat$pid3_num, useNA = "ifany")

# Race x Region interaction (16 categories)
# 1-4 = NE (White, Black, Hispanic, Other)
# 5-8 = MW (White, Black, Hispanic, Other)
# 9-12 = S (White, Black, Hispanic, Other)
# 13-16 = W (White, Black, Hispanic, Other)
dat <- dat %>%
  dplyr::mutate(
    race_eth_reg_num = dplyr::case_when(
      !is.na(race_num) & !is.na(region_num) ~ race_num + (region_num - 1) * 4,
      TRUE ~ NA_real_
    )
  )
table(dat$race_eth_reg_num, useNA = "ifany")
```

```{r Create factor variables for raking}
# Create _cat factor variables with meaningful text labels
# Keep the _num versions for other analyses
dat <- dat %>%
  dplyr::mutate(
    # Gender with labels
    gender_cat = factor(gender_num,
              levels = 1:2,
              labels = c("Male", "Female")),

    # Age with labels
    age_cat = factor(age_num,
              levels = 1:4,
              labels = c("18-29", "30-44", "45-64", "65+")),

    # Education with labels
    educ_cat = factor(educ_num,
              levels = 1:4,
              labels = c("HS or less", "Some college", "College", "Graduate")),

    # Income with labels
    inc5_cat = factor(inc5_num,
              levels = 1:5,
              labels = c("Up to $29,999", "$30,000-$59,999", "$60,000-$99,999",
                               "$100,000-$149,999", "$150,000+")),

    # Party ID with labels
    pid3_cat = factor(pid3_num,
              levels = 1:3,
              labels = c("Democrat", "Independent", "Republican")),

    # Race x Region with labels
    race_eth_reg_cat = factor(race_eth_reg_num,
              levels = 1:16,
              labels = c("NE White", "NE Black", "NE Hispanic", "NE Other",
                         "MW White", "MW Black", "MW Hispanic", "MW Other",
                         "S White", "S Black", "S Hispanic", "S Other",
                         "W White", "W Black", "W Hispanic", "W Other"))
  )

# Check factor levels with labels
table(dat$gender_cat, useNA = "ifany")
table(dat$age_cat, useNA = "ifany")
table(dat$educ_cat, useNA = "ifany")
table(dat$inc5_cat, useNA = "ifany")
table(dat$pid3_cat, useNA = "ifany")
table(dat$race_eth_reg_cat, useNA = "ifany")

# Completeness check for raking vars
dat <- dat %>%
  dplyr::mutate(
    rake_complete = !is.na(gender_num) &
      !is.na(age_num) &
      !is.na(educ_num) &
      !is.na(inc5_num) &
      !is.na(pid3_num) &
      !is.na(race_eth_reg_num)
  )

table(dat$rake_complete, useNA = "ifany")
nrow(dat)

# Filter to complete cases for raking
dat_complete <- dat %>% filter(rake_complete == TRUE)
nrow(dat_complete)
```

```{r Run survey rake}
# Create population margin data frames for survey::rake
# survey::rake needs data frames with the variable name and Freq column
# This is how many of each demographic category we'd expect to see in our sample if it was perfectly representative of the US population on that demographic

n <- nrow(dat_complete)

pop_gender <- data.frame(
  gender_cat = factor(names(target_gender), levels = levels(dat_complete$gender_cat)),
  Freq = as.numeric(target_gender) / 100 * n
)
pop_gender

pop_age <- data.frame(
  age_cat = factor(names(target_age), levels = levels(dat_complete$age_cat)),
  Freq = as.numeric(target_age) / 100 * n
)

pop_educ <- data.frame(
  educ_cat = factor(names(target_educ), levels = levels(dat_complete$educ_cat)),
  Freq = as.numeric(target_educ) / 100 * n
)

pop_inc5 <- data.frame(
  inc5_cat = factor(names(target_inc5), levels = levels(dat_complete$inc5_cat)),
  Freq = as.numeric(target_inc5) / 100 * n
)

pop_pid3 <- data.frame(
  pid3_cat = factor(names(target_pid3), levels = levels(dat_complete$pid3_cat)),
  Freq = as.numeric(target_pid3) / 100 * n
)

pop_race_eth_reg <- data.frame(
  race_eth_reg_cat = factor(names(target_race_eth_reg), levels = levels(dat_complete$race_eth_reg_cat)),
  Freq = as.numeric(target_race_eth_reg) / 100 * n
)

# Create unweighted survey design
unweighted_design <- svydesign(
  ids = ~1,
  data = dat_complete
)

# Rake the design
raked_design <- rake(
  design = unweighted_design,
  sample.margins = list(
    ~gender_cat,
    ~age_cat,
    ~educ_cat,
    ~inc5_cat,
    ~pid3_cat,
    ~race_eth_reg_cat
  ),
  population.margins = list(
    pop_gender,
    pop_age,
    pop_educ,
    pop_inc5,
    pop_pid3,
    pop_race_eth_reg
  ),
  control = list(maxit = 100, epsilon = 1e-4)
)

# Extract weights
dat_complete$wgtpi <- weights(raked_design)
summary(dat_complete$wgtpi)
hist(dat_complete$wgtpi, breaks = 50, main = "Distribution of Weights (wgtpi)")
```

```{r Trim weights}
# Cap = 7, floor = 0.2: No need to trim more aggressively like political polling. We're less concerned about a single point estimate of who's ahead and more concerned about a bunch of regression coefficients across different outcomes. We're also raking on more variables than in political polling, so weights are naturally going to inflate more, especially with the race x region interaction.
# A cap of 5 is more conservative, but 7 allows more correction for hard-to-reach demographic cells (e.g., young low-income Hispanic men in the South) without letting any single respondent dominate too heavily.

# Apply weight trimming
dat_complete <- dat_complete %>%
  dplyr::mutate(
    # Normalize to mean of 1 before trimming
    wgtpi = wgtpi / mean(wgtpi),
    # Cap at 7, floor at 0.2
    # Cap chosen to match colleague's Stata implementation; allows correction
    # for hard-to-reach demographic cells without any single respondent
    # dominating weighted estimates. A cap of 5 is more conservative.
    wgtpi = pmin(wgtpi, 7),
    wgtpi = pmax(wgtpi, 0.2)
    # Note: we do NOT re-normalize after trimming -- doing so would push
    # trimmed-at-7 cases above the cap. Weights are already close to mean 1
    # after the normalize-then-trim sequence.
  )

summary(dat_complete$wgtpi)
hist(dat_complete$wgtpi, breaks = 50, main = "Distribution of Weights (wgtpi)")

# Had previously tried iterative rake-trimming, but variance increased rather than decreased with each iteration. We're fine with one raking run and then trimming after in this case though. See code for iterative rake-trimming below:

# n_iter <- 3  # number of rake-trim cycles; 2-3 usually sufficient
#
# # Initialize equal weights before first iteration
# dat_complete$wgtpi <- 1
# 
# for (i in 1:n_iter) {
# 
#   raked_design <- rake(
#     design = svydesign(ids = ~1, data = dat_complete, weights = ~wgtpi),
#     sample.margins = list(
#       ~gender_cat, ~age_cat, ~educ_cat,
#       ~inc5_cat, ~pid3_cat, ~race_eth_reg_cat
#     ),
#     population.margins = list(
#       pop_gender, pop_age, pop_educ,
#       pop_inc5, pop_pid3, pop_race_eth_reg
#     ),
#     control = list(maxit = 100, epsilon = 1e-4)
#   )
# 
#   dat_complete$wgtpi <- weights(raked_design)
#   dat_complete$wgtpi <- dat_complete$wgtpi / mean(dat_complete$wgtpi)
# 
#   dat_complete$wgtpi <- pmin(dat_complete$wgtpi, 7)
#   dat_complete$wgtpi <- pmax(dat_complete$wgtpi, 0.2)
# 
#   cat(sprintf(
#     "Iteration %d — min: %.3f  max: %.3f  mean: %.3f  sd: %.3f\n",
#     i,
#     min(dat_complete$wgtpi),
#     max(dat_complete$wgtpi),
#     mean(dat_complete$wgtpi),
#     sd(dat_complete$wgtpi)
#   ))
# }
```

```{r Validate margins}
# Validate weighted margins vs targets
wprop <- function(x, w) {
  xt <- xtabs(w ~ x)
  as.numeric(xt) / sum(as.numeric(xt)) * 100
}

w_gender <- wprop(dat_complete$gender_num, dat_complete$wgtpi)
print(data.frame(
  category = c("Male", "Female"),
  target = as.numeric(target_gender),
  weighted = w_gender,
  diff = w_gender - as.numeric(target_gender)
))

w_age <- wprop(dat_complete$age_num, dat_complete$wgtpi)
print(data.frame(
  category = c("18-29", "30-44", "45-64", "65+"),
  target = as.numeric(target_age),
  weighted = w_age,
  diff = w_age - as.numeric(target_age)
))

w_educ <- wprop(dat_complete$educ_num, dat_complete$wgtpi)
print(data.frame(
  category = c("HS or less", "Some college", "College", "Graduate"),
  target = as.numeric(target_educ),
  weighted = w_educ,
  diff = w_educ - as.numeric(target_educ)
))

w_inc5 <- wprop(dat_complete$inc5_num, dat_complete$wgtpi)
print(data.frame(
  category = c("<$30k", "$30-60k", "$60-100k", "$100-150k", "$150k+"),
  target = as.numeric(target_inc5),
  weighted = w_inc5,
  diff = w_inc5 - as.numeric(target_inc5)
))

w_pid3 <- wprop(dat_complete$pid3_num, dat_complete$wgtpi)
print(data.frame(
  category = c("Democrat", "Independent", "Republican"),
  target = as.numeric(target_pid3),
  weighted = w_pid3,
  diff = w_pid3 - as.numeric(target_pid3)
))

w_race_eth_reg <- wprop(dat_complete$race_eth_reg_num, dat_complete$wgtpi)
race_eth_labels <- c(
  "NE White", "NE Black", "NE Hispanic", "NE Other",
  "MW White", "MW Black", "MW Hispanic", "MW Other",
  "S White", "S Black", "S Hispanic", "S Other",
  "W White", "W Black", "W Hispanic", "W Other"
)
print(data.frame(
  category = race_eth_labels,
  target = as.numeric(target_race_eth_reg),
  weighted = w_race_eth_reg,
  diff = w_race_eth_reg - as.numeric(target_race_eth_reg)
))
```

```{r Merge and save}
# Merge weights back to full dataset
dat_final <- dat %>%
  dplyr::left_join(
    dat_complete %>% dplyr::select(caseid, wgtpi),
    by = "caseid"
  )

nrow(dat_final)
# Weighted N
sum(!is.na(dat_final$wgtpi))
# Unweighted N
sum(is.na(dat_final$wgtpi))

# Save the weighted data
write.csv(dat_final, "data/data_weighted_r.csv", row.names = FALSE)
```

Comparison: Stata IPF vs. R survey::rake implementations
```{r Compare Stata IPF vs R rake implementations}
# Compares wgtpi from data_weighted_stata.csv (Stata IPF) vs dat_final (R survey::rake)

# Load data
stata_ipf <- read.csv("data/data_weighted_stata.csv", stringsAsFactors = FALSE)
r_rake <- dat_final

cat("Stata IPF N:           ", nrow(stata_ipf),                    "\n")
cat("R rake N:              ", nrow(r_rake),                       "\n")
cat("Stata IPF weighted N:  ", sum(!is.na(stata_ipf$wgtpi)),       "\n")
cat("R rake weighted N:     ", sum(!is.na(r_rake$wgtpi)),          "\n")

# Both Ns are the same, same people and same order. Can generate matching case IDs
stata_ipf$caseid <- seq_len(nrow(stata_ipf))
r_rake$caseid <- seq_len(nrow(r_rake))

# Weight distribution
cat("\n=== Weight Distribution ===\n")
cat("\nStata IPF weights:\n");  print(summary(stata_ipf$wgtpi))
cat("\nR rake weights:\n");     print(summary(r_rake$wgtpi))

par(mfrow = c(1, 2))
hist(stata_ipf$wgtpi, breaks = 50, main = "Stata IPF weights",  xlab = "wgtpi", col = "steelblue")
hist(r_rake$wgtpi,    breaks = 50, main = "R rake weights",     xlab = "wgtpi", col = "coral")
par(mfrow = c(1, 1))


# Targets as coded in Stata IPF do-file (line 175)
# ipfweight order: gender age educ inc5 pid3 race_eth_reg
# NOTE: Stata IPF targets differ from R rake targets on age, educ, income cutoffs, and race x region

stata_ipf_targets <- list(
  gender = c("Male" = 49.0, "Female" = 51.0),

  # Stata IPF age: 26.0 / 31.4 / 22.6 (vs. R rake 26.2 / 30.9 / 22.9)
  age = c("18-29" = 20.0, "30-44" = 26.0, "45-64" = 31.4, "65+" = 22.6),

  # Stata IPF educ: 37.5 / 29.0 / 20.7 / 12.8 (vs. R rake 37.3 / 28.5 / 21.0 / 13.2)
  educ = c("HS or less" = 37.5, "Some college" = 29.0, "College" = 20.7, "Graduate" = 12.8),

  # Stata IPF income brackets differ: $30k-$59,999 = codes 5-10, $60k-$99,999 = codes 11-18,
  # $100k-$149,999 = codes 19-20 only, $150k+ = codes 21-24 only (no "prefer not to answer")
  # targets: 17.4 / 20.5 / 38.6 / 23.6 / (last two not used here — see note)
  # Stata ipfweight has 5 income values: 18.4 20.6 22.1 17.4 21.5
  # Those come AFTER pid (45.9 7.4 46.7), cross-check the ordering in line 175:
  # gender(2) age(4) educ(4) inc5(5) pid3(3) race_eth_reg(16) = 34 values total
  # Line 175 values in order:
  #   gender:       49.0 51.0
  #   age:          20.0 26.0 31.4 22.6
  #   educ:         37.5 29.0 20.7 12.8
  #   inc5:         18.4 20.6 22.1 17.4 21.5
  #   pid3:         45.9 7.4 46.7
  #   race_eth_reg: 10.83 1.89 2.70 1.59 15.26 2.05 1.88 1.39 22.25 7.12 7.60 1.88 12.17 1.06 7.27 3.06
  inc5 = c("Up to $29,999" = 18.4, "$30,000-$59,999" = 20.6,
           "$60,000-$99,999" = 22.1, "$100,000-$149,999" = 17.4, "$150,000+" = 21.5),

  # Stata PID order: 1=Dem, 2=Ind, 3=Rep
  pid3 = c("Democrat" = 45.9, "Independent" = 7.4, "Republican" = 46.7),

  # Stata IPF race x region targets (from line 175, and comment on line 86)
  race_eth_reg = c(
    "NE White" = 10.83, "NE Black" = 1.89,  "NE Hispanic" = 2.70,  "NE Other" = 1.59,
    "MW White" = 15.26, "MW Black" = 2.05,  "MW Hispanic" = 1.88,  "MW Other" = 1.39,
    "S White"  = 22.25, "S Black"  = 7.12,  "S Hispanic"  = 7.60,  "S Other"  = 1.88,
    "W White"  = 12.17, "W Black"  = 1.06,  "W Hispanic"  = 7.27,  "W Other"  = 3.06
  )
)

r_rake_targets <- list(
  gender = c("Male" = 49.0, "Female" = 51.0),
  age    = c("18-29" = 20.0, "30-44" = 26.2, "45-64" = 30.9, "65+" = 22.9),
  educ   = c("HS or less" = 37.3, "Some college" = 28.5, "College" = 21.0, "Graduate" = 13.2),
  inc5   = c("Up to $29,999" = 16.8, "$30,000-$59,999" = 19.6,
             "$60,000-$99,999" = 20.8, "$100,000-$149,999" = 16.7, "$150,000+" = 26.1),
  pid3   = c("Democrat" = 45.0, "Independent" = 9.0, "Republican" = 46.0),
  race_eth_reg = c(
    "NE White" = 10.9,  "NE Black" = 1.78, "NE Hispanic" = 2.56,  "NE Other" = 2.04,
    "MW White" = 15.1,  "MW Black" = 1.92, "MW Hispanic" = 1.69,  "MW Other" = 1.65,
    "S White"  = 21.4,  "S Black"  = 6.82, "S Hispanic"  = 7.16,  "S Other"  = 3.33,
    "W White"  = 11.5,  "W Black"  = 1.02, "W Hispanic"  = 6.71,  "W Other"  = 4.3
  )
)

# Show target differences side by side
cat("\n=== Target Differences (Stata IPF vs. R rake) ===\n")
for (v in names(stata_ipf_targets)) {
  ipf <- stata_ipf_targets[[v]]
  rake <- r_rake_targets[[v]]
  out <- data.frame(
    category    = names(ipf),
    stata_ipf   = as.numeric(ipf),
    r_rake      = as.numeric(rake),
    diff        = round(as.numeric(rake) - as.numeric(ipf), 2)
  )
  cat("\n---", v, "---\n")
  print(out)
}


# Weighted margin validation (each vs. their own targets)
# Uses the variable names from R rake file; adjust stata_ipf names if CSV differs

wprop <- function(df, var, wgt = "wgtpi") {
  df <- df[!is.na(df[[wgt]]) & !is.na(df[[var]]), ]
  xt <- xtabs(as.formula(paste(wgt, "~", var)), data = df)
  round(as.numeric(xt) / sum(as.numeric(xt)) * 100, 2)
}

# Variable name mapping: list(var_in_stata_file = var_in_r_file)
# Adjust left-hand names if Stata CSV uses different column names
varmap <- list(
  gender_cat       = "gender_cat",
  age_cat          = "age_cat",
  educ_cat         = "educ_cat",
  inc5_cat         = "inc5_cat",
  pid3_cat         = "pid3_cat",
  race_eth_reg     = "race_eth_reg_cat"
)

ipf_target_list <- list(
  gender_cat       = stata_ipf_targets$gender,
  age_cat          = stata_ipf_targets$age,
  educ_cat         = stata_ipf_targets$educ,
  inc5_cat         = stata_ipf_targets$inc5,
  pid3_cat         = stata_ipf_targets$pid3,
  race_eth_reg     = stata_ipf_targets$race_eth_reg
)

cat("\n=== Weighted Margins: Stata IPF weights vs. Stata IPF targets ===\n")
for (var in names(varmap)) {
  if (!var %in% names(stata_ipf)) { cat("\nSkipping", var, "— not in Stata data\n"); next }
  w   <- wprop(stata_ipf, var)
  tgt <- as.numeric(ipf_target_list[[var]])
  cat("\n---", var, "---\n")
  print(data.frame(
    category = names(ipf_target_list[[var]]),
    target   = tgt,
    weighted = w,
    diff     = round(w - tgt, 2)
  ))
}

cat("\n=== Weighted Margins: R rake weights vs. R rake targets ===\n")
rake_target_list <- list(
  gender_cat       = r_rake_targets$gender,
  age_cat          = r_rake_targets$age,
  educ_cat         = r_rake_targets$educ,
  inc5_cat         = r_rake_targets$inc5,
  pid3_cat         = r_rake_targets$pid3,
  race_eth_reg     = r_rake_targets$race_eth_reg
)
for (var in names(varmap)) {
  var_r <- varmap[[var]]
  if (!var_r %in% names(r_rake)) { cat("\nSkipping", var_r, "— not in R data\n"); next }
  w   <- wprop(r_rake, var_r)
  tgt <- as.numeric(rake_target_list[[var]])
  cat("\n---", var_r, "---\n")
  print(data.frame(
    category = names(rake_target_list[[var]]),
    target   = tgt,
    weighted = w,
    diff     = round(w - tgt, 2)
  ))
}


# Individual weight correlation (matched on caseid)
if ("caseid" %in% names(stata_ipf) & "caseid" %in% names(r_rake)) {
  merged <- inner_join(
    stata_ipf %>% select(caseid, wgtpi) %>% rename(wgtpi_stata = wgtpi),
    r_rake    %>% select(caseid, wgtpi) %>% rename(wgtpi_r     = wgtpi),
    by = "caseid"
  ) %>% filter(!is.na(wgtpi_stata) & !is.na(wgtpi_r))

  cat("\n=== Individual Weight Correlation ===\n")
  cat("N matched cases:         ", nrow(merged), "\n")
  cat("Pearson r:               ", round(cor(merged$wgtpi_stata, merged$wgtpi_r), 4), "\n")
  cat("Mean absolute difference:", round(mean(abs(merged$wgtpi_stata - merged$wgtpi_r)), 4), "\n")
  cat("Max absolute difference: ", round(max( abs(merged$wgtpi_stata - merged$wgtpi_r)), 4), "\n")

  plot(merged$wgtpi_stata, merged$wgtpi_r,
       xlab = "Stata IPF weights", ylab = "R rake weights",
       main = "Weight correlation", pch = 16, col = rgb(0, 0, 0, 0.2))
  abline(0, 1, col = "red", lty = 2)

} else {
  cat("\nNo shared 'caseid' found — skipping individual correlation.\n")
}

# Very close -- r = .98
# However there is a max difference of ~3.3; worth checking
merged %>%
  mutate(abs_diff = abs(wgtpi_stata - wgtpi_r)) %>%
  arrange(desc(abs_diff)) %>%
  head(20) %>%
  left_join(
    r_rake %>% select(caseid, gender_cat, age_cat, educ_cat, inc5_cat, pid3_cat, race_eth_reg_cat),
    by = "caseid"
  )
# virtually every high-divergence case is S Other + $150k+, with one W Other + $150k+. These are the two cells where R rake targets diverge most from Stata IPF (S Other: 3.33 vs. 1.88; $150k+: 26.1 vs. 21.5), and the raking algorithm is compounding those differences for respondents who fall in both simultaneously. Nothing necessarily wrong with either set

# Substantive outcome spot-check
outcomes <- c("Q220_1", "Q220_2", "Q78")

cat("\n=== Weighted Outcome Means (spot check) ===\n")
cat(sprintf("%-10s  %9s  %9s  %8s\n", "variable", "Stata_IPF", "R_rake", "diff"))
for (v in outcomes) {
  if (!v %in% names(stata_ipf) | !v %in% names(r_rake)) next
  m_ipf  <- weighted.mean(stata_ipf[[v]], stata_ipf$wgtpi, na.rm = TRUE)
  m_rake <- weighted.mean(r_rake[[v]],    r_rake$wgtpi,    na.rm = TRUE)
  cat(sprintf("%-10s  %9.3f  %9.3f  %8.3f\n", v, m_ipf, m_rake, m_rake - m_ipf))
}
# weights are working correctly. Both implementations hit their respective targets closely, and the outcome means (Q220_1, Q220_2, Q78) are essentially identical across the two weighting schemes — differences of 0.03 or less. That's strong evidence the two approaches are producing substantively equivalent results despite the methodological differences (IPF in Stata vs. survey::rake in R).
```

Survey Codebook

[Age; Q2] What is your age?
•	Younger than 18 (1, excluded)
•	18–29 (2)
•	30–44 (3)
•	45–64 (4)
•	65 or older (5)

[Gender, Q3] Are you…
•	Female (1)
•	Male (2)

[Hispanic Ethnicity, Q4] Do you identify as Hispanic or Latino?
•	Yes (1)
•	No (2)

[Race, Q5] Which racial group describes you the most?
•	White (1)
•	Black (2)
•	Asian (3)
•	Native American (4)
•	Pacific Islander (5)
•	Middle Eastern or North African (6)
Race and ethnicity were recoded into four mutually exclusive dummy variables for analysis. Respondents who answered “yes” to the Hispanic ethnicity question were coded as Hispanic regardless of the racial category they selected. Among respondents who indicated they were not Hispanic, those who selected White were coded as non-Hispanic White, those who selected Black were coded as non-Hispanic Black, and all remaining racial selections (including Asian, Native American, Pacific Islander, Middle Eastern or North African, or any other category) were coded as Other race.

[Party Identification, Q10] Do you consider yourself a Democrat, a Republican, or an Independent?
•	Democrat (1)
•	Republican (2)
•	Independent (3)

[Party Lean, Q13] Do you consider yourself to be leaning Democrat, leaning Republican or neither?
•	Leaning Democrat (1)
•	Leaning Republican (2)
•	Neither (3)
